\documentclass{exam}
\usepackage[utf8]{inputenc}
\usepackage{geometry,dirtytalk,amsmath}
\usepackage{draftwatermark}
\usetikzlibrary{scopes}
\SetWatermarkText{\textit{Barron's High Frequency 333}}
\begin{document}
\begin{center}
\begin{huge}
Barron's 333 High Frequency
\end{huge}
\end{center}
\begin{questions}
\question \textbf{Abate} - $v.$ to subside; to reduce,  mitigate.\\
$\cdot$ Rather than leaving immediately, they waited for the storm to $abate$.
\question \textbf{Aberrant} - $adj.$ abnormal or deviant,  antic, anomalous, preternatural, atypical, grotesque, bizarre. \\
$\cdot$ Given the $aberrant$ nature of the data, we came to doubt the validity of the entire experiment.
\question \textbf{ Abeyance} - $n.$ suspended action, hiatus\\
$\cdot$ The deal was held in $abeyance$ until her arrival. \\
 $\cdot$ Expansion plans for the middle school are in $abeyance.$
\question \textbf{ Abscond} - $v.$ depart secretly and hide.\\
$\cdot$ The teller $absconded$ with bonds went uncaptured until someone recognized him from his photograph on $America's\,Most\,Wanted.$
\question \textbf{Abstemious} - $adj.$ sparing in eating and drinking; temperate, sober, austere, ascetic, puritanical, spartan.\\
$\cdot$ Concerned whether her vegetarian son's $abstemious$ diet provided him sufficient protein, the worried mother pressed food on him.\\
$\cdot$ The monks here have willingly chosen this $abstemious$ life.
\question \textbf{Admonish} - $v.$ warn; reprove, 	reprimand, rebuke, rant, rail, scold, reproach, upbraid, chastise, chide, berate, criticize, browbeat. \\
$\cdot$ He $admonished$ his listeners to change their wicked ways. \textbf{Admonition} $n.$\\
$\cdot$ She $admonished$ him to drink less.
\question \textbf{Adulterate} - $v.$ make impure by adding inferior or tainted substances. Violate and sanctity, besmirch, defile, desecrate, taint, profane, sully.\\
$\cdot$ It is a crime to $adulterate$ foods without informing the buyer; when consumers learn that Beechnut had $adulterated$ it's apple juice by mixing the juice with water, they protested vigorously. \textbf{Adultration} - $n$.
\question \textbf{Aesthetic} - $adj.$ artistic; dealing with or capable of appreciating the beautiful. Pulchritude.\\
$\cdot$ The beauty of Tiffany's stained glass appealed to Alice's $asthetic$ sense. \textbf{Aesthete}- $n.$
\question \textbf{Aggregate} - $v.$ gather; accumulate.\\
$\cdot$ Before the Wall Street scandals, dealers in so-called junk bonds manage to $aggregate$ great wealth in short period of time.
\question \textbf{ Alacrity} - $n.$ cheerful promptness; eagerness, Effervescence, Exuberance, Gusto, Verve, Furor, Vim, Zeal.\\
$\cdot$ Phil and Dave were raring to get off to the mountains; they packed up their ski gear and climbed into the van with $alacrity.$\\
$\cdot$ She accepted the invitation with $alacrity.$
\question \textbf{Alleviate} - $v.$ relieve. Allay, Appease, Assuage, Condole, Console, Mollify, Pacify, Concialiate, Placate, Propitiate.\\
$\cdot$ This should $alleviate$ the pain; if it does not we shall have to use stronger drugs.
\question \textbf{Amalgamate} - $v.$ combine; unite in one body.\\
 $\cdot$ The unions will attempt to $amalgamate$ their groups into one national body.
 \question \textbf{Ambiguous} - $adj.$ unclear or doubtful in meaning.\\
 $\cdot$ His $ambiguous$ instruction misled us; we didn't know which road to take. \textbf{ambiguity-$n.$}\\
 \question \textbf{Ambivalence} - $n.$ the state of having contradictory or conflicting emotional attitudes.\\
 $\cdot$ Torn between loving her parents one minute and hating them the next, she was confused the ambivalence of her feelings. \textbf{Ambivalent - $adj.$}
\question \textbf{Ameliorate} - $v.$ improve\\
 $\cdot$ Many social workers are attempted to ameliorate the conditions of living in the slums.
\question \textbf{Anachronism} - $n.$ something or someone misplaced in time\\
 $\cdot$ Shakespeare's reference to clocks in $Julius\, Caesar$ is an $anachronism$; no clocks existence in Caesar's time.
 \question \textbf{Analogous} - $adj.$  comparable.\\
$\cdot$ She called our attention to think that had been done in an $analogous$ situation and recommended that we do the same. 
 \question \textbf{Anarchy} - $n.$ absence of governing body; state of disorder\\
$\cdot$ The assassination of the leaders led to the period of $anarchy$.
 \question \textbf{Anomalous} - $adj.$  abnormal; irregular.\\
$\cdot$ She was placed to $anomalous$ position of seeming to approved procedures that she despised.  
 \question \textbf{Antipathy} - $n.$  aversion; dislike.\\
$\cdot$ Tom's extreme $antipathy$ for disputes keeps him from getting into arguments with his temperamental wife. Noise in any form is $antipathetic$ to him. Among his other $antipathies$ are honking cars, boom boxes and heavy metal rock.
 \question \textbf{Apathy} - $n.$  lack of caring; indifference.\\
$\cdot$ A firm believer in democratic government, she couldn't understand the $apathy$ of people who never bothered to vote. \textbf{Apathetic - $adj.$}
 \question \textbf{Appease} - $adj.$  pacify or soothe; relieve.\\
$\cdot$ Tom and Jody tried to $appease$ the crying baby by offering him toy one after another. However he couldn't calm down until they $appeased$ his hunger by giving him a bottle. \textbf{Appeasement- $n.$} 
 \question \textbf{Apprise} - $adj.$  inform.\\
$\cdot$ When she was $apprised$ of the dangerous weather conditions, she decided to postponed her trip.  
 \question \textbf{Approbation} - $adj.$  approval.\\
$\cdot$ Wanting her parents regard, she looked for some sign of their $approbation$.
 \question \textbf{Appropriate} - $adj.$  acquire; take possession of for one's own use.\\
$\cdot$ The ranch owners $appropriated$ the lands that had originally been set aside for the Indian's use.
 \question \textbf{Arduous} - $adj.$  hard; strenuous; onerous; Herculean.\\
$\cdot$ Her $arduous$ effort had sapped her energy.
 \question \textbf{Artless} - $adj.$  without guile; open and honest.\\
$\cdot$ Red Riding Hood's $artless$ comment \say{Grandma, What big eyes you have!} indicates the child 's innocent surprise at her \say{grandmother's} changed appearance. 
 \question \textbf{Ascetic} - $adj.$  practicing self-denial; austere.\\
$\cdot$ The wealthy shelf-indulgent young man oddly drawn to the strict, ascetic life led by the members of some monastic orders. also \textbf{asceticism-$n.$} 
 \question \textbf{Assiduous} - $adj.$ Careful; Meticulous; Punctilious; Scrupulous; Sedulous; Demure; Diligent; Circumspect; Painstaking; Politic; Prudent; Wary; Chary; Leery.\\
$\cdot$ It took Rembrandt weeks of $assiduous$ labour before he was satisfied with his portrait of his son. 
 \question \textbf{Assuage} - $adj.$ ease or lessen (pain); satisfy (hunger); soothe (anger); Calm, Mollify, Pacify; Alleviate; Appease; Condole; Console; Conciliate; Placate; Propitiate.\\
$\cdot$ Dick tried to $assuage$ heartache by indulging in ice cream. One gallon later, he has $assuaged$ his appetite but not his grief. \textbf{assuagement-$n.$}
 \question \textbf{Attenuate} - $adj.$  make thinner; Ebb; Damp; Alloy.\\
$\cdot$ By withdrawing their forces, the generals hoped to $attenuate$ the enemy lines.
 \question \textbf{Audacious} - $adj.$  daring; bold.\\
$\cdot$ Audiences cheered as Luke Skywalker ans princes Leia made their $audacious$, death-defying leap to freedom and escape Darth Vader's troops. \textbf{audacity-$n.$}
 \question \textbf{Austere} - $adj.$ forbiddingly stern; severely simple and unornamented.\\
$\cdot$ The Head Master's $austere$ demeanor tended to scare off the more timid students, who never visited his study willingly. The room reflected the man austere and bare, like a monk's cell, with touches of luxury to moderate its $austerity$.
 \question \textbf{Autonomous} - $adj.$  self-governing; independent.\\
$\cdot$ Although the University of California at Berkeley is just a part of state university system Cal Berkeley is $autonomous$, for its runs several programs that are not subject to out side control. \textbf{autonomy-$n.$}
 \question \textbf{Aver} - $adj.$  assert confidently or declare; as used in law, state formally as a fact.\\
$\cdot$ I wish to $aver$ I am certain of success.
 \question \textbf{Banal} - $adj.$ hackneyed; commonplace; trite; lacking originality.\\ 
 $\cdot$ The hack writer's worn-out clinches made his comic sketch seem $banal$. He even restored the $banality$ of having someone sleep on a banana peel.. 
  \question \textbf{Belie} - $adj.$  contradict; give a false impression\\
 $\cdot$ His coarse, hard-bitten exterior $belied$ his innate sensitivity.
  \question \textbf{Beneficent} - $adj.$ kindly; doing good\\
 $\cdot$ The over generous philanthropist had to curb his $beneficent$ impulses before he gave away all his money and left himself with nothing. 
  \question \textbf{Bolster} - $adj.$  support; reinforce.\\
 $\cdot$ The debaters amassed file boxes full of evidence to bolster their arguments.
  \question \textbf{Bombastic} - $adj.$  pompous; using inflated language
 $\cdot$
  \question \textbf{Aver} - $adj.$  assert 
 $\cdot$
  \question \textbf{Aver} - $adj.$  assert 
 $\cdot$
  \question \textbf{Aver} - $adj.$  assert 
 $\cdot$
\end{questions}
\end{document}