\documentclass{exam}
\usepackage[utf8]{inputenc}
\usepackage{geometry,dirtytalk,amsmath}
\usepackage{tikz}
\usepackage{draftwatermark, tcolorbox}
\usetikzlibrary{scopes}
\SetWatermarkText{\textit{Direct Hits Core Vocabulary }}
\begin{document}
\begin{center}
\begin{huge}
Direct Hits Core Vocabulary
\end{huge}
\end{center}
\begin{questions}
\question  \textbf{AMBIVALENT}\\
\textit{Contradictory, having mixed feelings.}\\

$\cdot$ In the TV show Glee, Finn Hudson is the star
quarterback of his high school football team. Finn is
also a talented singer who is AMBIVALENT about
joining the school Glee Club. Although Finn doesn't
want to alienate his teammates, he enjoys singing and
wants to follow his dream of becoming an entertainer.\\

$\cdot$ In the movie The Notebook, Allie has to choose
between Noah and Lon. She is emotionally torn by her
\textbf{AMBIVALENT} feelings as she tells Noah, \say{There is no
easy way, no matter what I do, somebody gets hurt}.
She later reiterates her \textbf{AMBIVALENT} feelings when
she tells Lon \say{When I'm with Noah I feel like one
person, and when I'm with you, 1 feel like someone
totally different}.

\begin{tcolorbox}
\begin{center}
\textbf{ KNOW YOUR ROOTS}\\
\textbf{LATIN PREFIX: AMBI } - \textit{Both}.\\  
\end{center}  
\textbf{AMBIDEXTROUS} - \textit{able to use both hands with equal ease, skillful, versatile}\\
\textbf{AMBIGUOUS} (Word 21) \textit{having two or more possible meanings, doubtful, dubious,} \textbf{ EQUIVOCAL} (Word 210)\\
\textbf{AMBIVALENT} - \textit{ being simultaneously of two mind}
\end{tcolorbox}

\question  \textbf{ANOMALY}\\ \textit{Deviation from the norm, something that is}
\textbf{ATYPICAL}.\\

$\cdot$ In the Harry Potter series, Vernon Dursley prides
himself on being \say{perfectly normal, thank you very
much}. An \textbf{ANOMALY} is the last thing Dursley wants
in his life. In the opening chapter of Harry Potter and
the Sorcerer's Stone, however, Dursley notices several
strange \textbf{ANOMALIES} or what he calls \say{funny stuff}.
For example, be spots a cat that appears to be reading
a map. He also notices a number of odd people who
are dressed in colorful robes. And that is not all: flocks
of owls can be seen flying during the daytime.

\question \textbf{SARCASTIC, SARDONIC, SNIDE}\\ \textit{Mocking, derisive, taunting, and stinging.}\\

$\cdot$ Winston Churchill was famous for his \textbf{SARCASTIC}
and \textbf{SARDONIC} comments. Here are two well-known
examples:\\
Bessie Braddock: Sir, you are a drunk.\\
Churchill: Madame, you are ugly. In the morning I shall be sober, and you will still be ugly.\\
Nancy Astor·: Sir, if you were my husband I would give you poison.\\
Churchill: If I were your husband I would take it.\\
In the movie Avatar, Dr. Grace Augustine tells Jake,
\say{Just relax and let your mind go blank. That shouldn't
be too hard for you}. This \textbf{SNIDE} remark expresses
Grace's initial contempt for Jake.

\question \textbf{DEARTH and PAUCITY}\\ \textit{A scarcity or shortage of something.}\\

$\cdot$ A study by USA Today revealed that college football's top level teams have a \textbf{DEARTH} of minority coaches. Just 10 of the 120 Football Bowl Subdivision
universities have minority head coaches. There is also
a \textbf{PAUCITY} of minority assistant coaches. In contrast,
about one-fourth of the 32 NFL teams have minority
head coaches. Critics encourage the colleges to
\textbf{EMULATE} (imitate) the NFL's policy of requiring
teams to interview minority candidates for head
coaching opportunities.

\question \textbf{PRATTLE}\\ \textit{To speak in a foolish manner, to babble incessantly.}\\

$\cdot$ In the movie Office Space, Milton continuously
\textbf{PRATTLES} to himself about how he is abused by
management and bow his co-workers repeatedly
borrow his stapler. Here is an example of Milton
\textbf{PRATTLING} on and on:\\
\say{I don't care if they Lay me off either, because I
told, I told Bill that if they move my desk one
more time, then, then, I'm, I'm quitting, I'm
going to quit. And, I told Don too, because
they've moved my desk four times already this
year, and I used to be over by the window, and
I could see the squirrels, and they were
married, but then, they switched from the
Swingline to the Boston stapler, but I kept my
Swingline stapler because it didn't bind up as
much, and I kept the staplers for the Swingline
stapler and it's not okay, because if they take
my stapler then I'll set the building on fire $\cdots$}

\question \textbf{WRY} \\
\textit{Dry, humorous with a clever twist and a touch of irony;} \textbf{DROLL}\\

$\cdot$ Casel Di Principe is normally a sleepy town northwest of Naples, Italy. On a quiet Sunday a 6o-year-old man sat at a table playing cards -with his friends. Two
gunmen suddenly shattered the peace when they
drove down the town's main street and fired a barrage
of bullets, killing the old man. The assassination of an
important crime boss caused authorities to launch a
full-scale investigation. However, the witnesses refused to answer questions, claiming that they didn't see anything. Their collective silence prompted a \textbf{WRY} comment from one police officer that \say{the victim must
have been playing solitaire}.\\

\begin{tcolorbox}
A WRY sense of humor is very different from a
\textbf{JOCULAR} sense of humor. A \textbf{WRY} joke appeals
to your intellect and often produces a knowing
smile. In contrast, a \textbf{JOCULAR} joke appeals to
your funny bone and produces a belly laugh.
\end{tcolorbox}

\question \textbf{UNCONVENTIONAL} and
\textbf{UNORTHODOX}\\ \textit{Not ordinary or typical; characterized by
avoiding customary conventions and behaviors.}\\

$\cdot$ Both Katy Perry and Lady Gaga are known for their
catchy hits and bold, \textbf{UNCONVENTIONAL}
wardrobes. Katy Perry's colorful, \textbf{UNORTHODOX}
costumes include a funky playing cards dress, a
watermelon dress, and a cute pink panda outfit.\\

Lady Gaga is also known for wearing
\textbf{UNCONVENTIONAL} and even \textbf{OUTLANDISH}
(bizarre, outrageous) stage outfits. Gaga is usually
attired in her trademark platinum blonde hair, funky
sunglasses, and revealing leotards. However, she is
always unpredictable. Gaga opened one concert
wearing clear plastic champagne bubbles and then
changed into a sculpted prom dress.

\question \textbf{PAINSTAKING} and \textbf{METICULOUS}\\ \textit{Extremely careful; very} \textbf{EXACTING}\\

$\cdot$ The new Harry Potter theme park is a \textbf{METICULOUS}
recreation of Hogwarts castle and nearby Hogsmeade
village. The park's designers spared no expense to
\textbf{PAINSTAKINGLY} recreate such iconic rooms as
Dumbledore's office and the Defense Against the Dark
Arts classroom. \textbf{ENTHRALLED} (fascinated) visitors
can sample butterbeer and even purchase a wand at
Ollivander's Wand Shop.

\question \textbf{AUDACIOUS}\\
\textit{Fearlessly, often recklessly daring; very bold.}\\

$\cdot$ What do American General George Washington and
Japanese Admiral Isoroku Yamamoto have in
common? Both launched \textbf{AUDACIOUS} surprise
attacks on unsuspecting adversaries. On Christmas
Day, 1776, Washington ordered the Colonial Army to
cross the Delaware and attack the British and Hessian
forces at Trenton. Washington's \textbf{AUDACIOUS} plan
shocked the British and restored American morale.
On December 7, 1941, Yamamoto ordered the
Japanese First Air Fleet to launch a surprise attack on
the American Pacific Fleet based at Pearl Harbor.
Although Japan's \textbf{AUDACIOUS} sneak attack
temporarily \textbf{HOBBLED} (hampered) the U.S. fleet, it
aroused the now unified country to demand revenge.

\question  \textbf{INDIFFERENT and APATHETIC}\\ \textit{Marked by a lack of interest or concern.}\\

$\cdot$ In the movie Ferris Bueller's Day Off, the economic
teacher Ben Stein delivers a \textbf{SOPORIFIC} (sleep inducing) lecture on tariffs and the Great Depression.
Stein's bored and \textbf{INDIFFERENT} students ignore his
monotone lecture. Hoping for some sign of interest,
Stein tries asking questions, but his efforts are
\textbf{FUTILE} (Word 46). Some students are so
\textbf{APATHETIC} they fall asleep.

\question \textbf{DIFFIDENT}\\
\textit{Lacking self-confidence; self-effacing; NOT assertive.}\\

$\cdot$ Even Lady Gaga wakes up and feels like an insecure and \textbf{DIFFIDENT} 24-year-old girl. But Gaga then tells herself, \say{You're Lady Gaga, you get up and walk the walk today.} As you study for the SAT, be like Lady
Gaga. Don't \textbf{SUCCUMB} (give in) to feelings of
\textbf{DIFFIDENCE}. Study your Direct Hits vocabulary and
be confident.

\question \textbf{PRAGMATIC}\\ \textit{Practical; sensible; NOT idealistic or romantic.}\\

$\cdot$ What do the 16th century French king Henry N and
the 20th century American president Franklin Delano
Roosevelt have in common? Both leaders made
PRAGMATIC decisions that helped resolve a crisis.
Henry IV was the newly crowned Protestant king in a
country dominated by Catholics. For the sake of his
war-weary country, Henry IV \textbf{PRAGMATICALLY}
chose to become a Catholic, saying, \say{Paris is worth a Mass}.\\
FDR was a newly-elected president in a country facing
the worst economic crisis in its history. For the sake of
his country, Roosevelt \textbf{PRAGMATICALLY} chose to replace traditional laissez-faire economic policies with
\say{bold, persistent experimentation}. FDR \textbf{PRAGMATICALLY} explained, \say{It is common sense to take a method and try it; if it fails, admit it frankly and try
another. But above all, try something}.

\question  \textbf{EVOCATION}\\
\textit{An imaginative re-creation.}\\

$ \cdot $ What do the treasures of Pharaoh Tutankhamen,
Taylor Swift's music video \say{Love Story}, and the
movie Titanic all have in common? They are all
powerful \textbf{EVOCATIONS}. The treasures of Pharaoh
Tutankhamen are \textbf{EVOCATIONS} of the power and
splendor of Ancient Egypt. Taylor Swift's \say{Love Stacy}
\textbf{EVOKES} a time when beautiful princesses lived in
romantic castles and fell in love with handsome
princes. And the movie Titanic is a remarkable
\textbf{EVOCATION} of what it was like to be a passenger on
the great but doomed ship.

\begin{tcolorbox}
\begin{center}
\textbf{ KNOW YOUR ROOTS}\\
\textbf{LATIN ROOT: VOC, VOK} - \textit{Call}.\\  
\end{center}  
\textbf{VOCAL} - \textit{related to the voice, speaking}\\
\textbf{VOCATION} - \textit{your calling, your profession, often used for a religious career}\\
\textbf{AVOCATION} -  \textit{a second calling, a hobby}\\ 
\textbf{EVOKE} -  \textit{to call forth, especially from the past}\\
\textbf{REVOKE} -  \textit{to call back, to rescind, to repeal}\\ 
\textbf{INVOKE} - \textit{to call upon. Epic poems often begin with an Invocation of the Muse, or goddess of artistic inspiration}\\
\textbf{PROVOKE} -  \textit{to call forth} (see Word 82)\\ 
\textbf{CONVOCATION} - \textit{a calling together, a gathering}\\
\textbf{VOCIFEROUS} - \textbf{making an outcry, clamorous}\\
\textbf{EQUIVOCATE} -  \textit{to use} \textbf{AMBIGUOUS} (Word 21) \textit{expressions, to mislead} \textbf{IRREVOCABLE} \textit{incapable of being recalled or altered.}
\end{tcolorbox}

\question \textbf{PRESUMPTUOUS}\\
\textit{Overbearing; impertinently bold; characterized by brashly overstepping one's place.}\\

$ \cdot $ In the movie 300, Queen Gorgo boldly told the
Persian envoy, \say{Do not be coy or stupid Persian. You
can afford neither in Sparta.} Queen Gorgo's willingness to speak out astonished the Persian envoy.
Shocked by the Spartan Queen's \textbf{PRESUMPTUOUS}
statement, the envoy questioned, \say{What makes this
woman think she can speak among men?}

\question \textbf{RECALCITRANT}\\
\textit{Stubborn resistance to and defiance of authority or guidance;} \textbf{OBSTINATE; OBDURATE.}\\

$ \cdot $ What do Hester Prynne (The Scarlet Letter) and the
singer Amy Winehouse have in common? Both are
\textbf{RECALCITRANT}. In The Scarlet Letter, the Reverend
Wilson demanded that Hester reveal the name of the
father of her child. But Hester was \textbf{RECALCITRANT}.
Despite \say{the heavy weight of a thousand eyes, all
fastened upon her,} Hester stubbornly refused to
name the father, defiantly declaring, \say{Never...! will
not speak!} In her song \say{Rehab,} Amy Winehouse is
also defiantly \textbf{RECALCITRANT}. Her friends and
family all beg her to go to rehab, but Amy is
\textbf{OBDURATE} and defiantly declares, \say{No, no, no.}

\question \textbf{BOON}\\ \textit{A timely benefit; blessing.}\\ \textbf{Bane}\\
\textit{A source of harm and ruin.} \\

$ \cdot $ Fifty Cent was shot nine times and lived! Was the
shooting a \textbf{BANE} or a \textbf{BOON} for his career? At first it
was a \textbf{BANE} because the pain was excruciating and
Fiddy bad to spend weeks in a hospital recuperating.
But the shooting turned out to be a \textbf{BOON} for his
career because it \textbf{BOLSTERED} (reinforced) Fiddy's
\say{street cred} and attracted lots of publicity.

\question \textbf{CLANDESTINE} and \textbf{SURREPTITIOUS }\\
\textit{Secret; covert; not open;} \textbf{NOT ABOVEBOARD}.\\

$ \cdot $ What do the Men in Black (Men In Black),
Dumbledore's Army (Harry Potter and the Order of the Phoenix, and Sector Seven (Transformers) all
have in common? They are all \textbf{CLANDESTINE} groups
that conduct \textbf{SURREPTITIOUS} activities. The Men in Black \textbf{SURREPTITIOUSLY} regulate alien life forms on Earth. Dumbledore's Army teaches Hogwarts students how to defend themselves agajnst the Dark Arts.
And Sector Seven guards the mysterious All Spark and
keeps the body of Megatron permanently frozen.
\question \textbf{AFFABLE, AMIABLE, GENIAL,
GREGARIOUS}\\ \textit{All mean agreeable; marked by a pleasing
personality; warm and friendly.}\\

$ \cdot $ President Reagan was renowned for his \textbf{AFFABLE}
grace and \textbf{GENIAL} good humor. On March 6, 1981, a
deranged gunman shot the president as he was
leaving a Washington hotel. The injured but always
\textbf{AMIABLE} president looked up at his doctors and
nurses and said, \say{I hope you're all Republicans.} The
first words be uttered upon regaining consciousness
were to a nurse who happened to be holcling the
president's hand. \say{Does Nancy know about us?} the
president joked.

\begin{tcolorbox}

If you think you have heard the word \textbf{AMIABLE} before, you are probably right. The English word
\textbf{AMIABLE} contains the Latin root ami meaning
friend. You may have beard this root in the
French word ami and the Spanish word amigo.
\begin{center}
\textbf{ KNOW YOUR ROOTS}\\
\textbf{LATIN PREFIX: AMI} - \textit{Friend}.\\  
\end{center}  
\textbf{AMITY} - \textit{Friendship, harmony}\\
\textbf{AMICABLE} - \textit{peaceable, harmonious}
\end{tcolorbox}

\question \textbf{AUSTERE}\\
\textit{Having no adornment or ornamentation; bare;
not} \textbf{ORNATE} (Word 363).\\ \textbf{AUSTERITY}\\ \textit{The trait of great self-denial; economy.}\\

$ \cdot $ Ancient Greek architects often used Doric columns to construct temples. For example, the Parthenon's \textbf{AUSTERE} columns conveyed strength and simplicity because they lacked ornamentation.\\

Although modern Greeks admire the \textbf{AUSTERE}
columns built by their ancestors, they vigorously
oppose new \textbf{AUSTERITY} measures that raise taxes
and cut social welfare programs. These \textbf{AUSTERITY}
measures have provoked massive protests.

\question \textbf{ALTRUISTIC}\\\textit{Unselfish concern for the welfare of others.}\\

$ \cdot $ Eleven-year-old Olivia Bouler sobbed uncontrollably
when she first saw pictures of oil-coated birds dying
along the Gulf coast. Rather than continue to cry,
Olivia wrote a Jetter to the Audubon Society offering
to help. Olivia volunteered to draw and sell watercolor
paintings o fbirds and give the profits to the Audubon
Society. Olivia's \textbf{ALTRUISTIC} campaign has helped
raise over \$ 160,000 to save Gulfbirds.

\question \textbf{AMBIGUOUS}\\
\textit{Unclear; uncertain; open to more than one interpretation; not definitive.}\\

$ \cdot $ Tbe final scene of the movie Inception is deliberately \textbf{AMBIGUOUS}. Leo DiCaprio's character, Dom Cobb,is \textbf{ELATED} (very happy) because he has found his
children and completed the seemingly impossible job
he was hired to do. But is all this real or is Dom
entrapped in yet another dream? Dom uses a metal
top to enable him to determine what is real and what
isn't. At the end of the film, Dom spins the top. What
will happen next? If the top keeps spinning, Dom is
dreaming. If it falls, things are real. We don't know
what happens because the ending is \textbf{AMBIGUOUS}.

\question  \textbf{UPBRAID, REPROACH, CASTIGATE}\\ \textit{To express disapproval; to scold; to rebuke; to }\textbf{CENSURE.}\\

$\cdot$ In this classic scene from Billy Madison, Ms. Vaughn \textbf{UPBRAIDS} Billy for making fun of a third grade student who is having trouble reading: \\Third Grader: Wa-wa-wa-once th-th-th-th-there wa-wa-wa-was a-a-a-a g-g-girl \\Billy Madison: Kid can't even read. \\Ernie: Cut it out, dude, you're gonna get us in trouble. \\Billy Madison: T-T-T-Today Junior! \\Billy Madison: OW! You're tearing my ear off! \\Veronica Vaughn : Making fun of a little kid for trying to read. Are you psycho? Do you not have a soul? You keep your mouth shut for the next two weeks or I'm going to fail you. End of story.

\question  \textbf{Nostalgia}\\
\textit{A \textbf{WISTFUL} (Word 206) sentimental longing for a place or time in the past.}\\

$\cdot$ A lifelong fan of The Muppets, Jason Segel was \textbf{NOSTALGIC} for his childhood, and he decided to  \textbf{REJUVENATE} (Word 171) the franchise by writing a new movie for them. Segel said, \say{We set out to make a Muppet movie that harkened back to the late-'70s, early-'80s Muppets that we grew up with.} It's been over a decade since The Muppets starred in a theatrical movie, and, likewise, in The Muppets, it's been a while since Kermit and his friends have performed as a group. As the audience revisits their childhood icons during this \textbf{NOSTALGIC} film, The Muppets, too, take a \textbf{WISTFUL} (Word 206) walk down memory lane. The Muppets decide to get their group together again for one last show, but they discover that they aren't popular anymore. They have become \textbf{ANTIQUATED} (Word 25); one character tells them, \say{You're \textbf{RELICS} (surviving objects from the past).} By incorporating clever humor and  \textbf{WISTFUL} (Word 206) references to Muppet movies of the past, The Muppets introduces a new generation to the \textbf{WHIMSICAL} (Word 219) world of Kermit and his friends while also catering to an older \textbf{DEMOGRAPHIC's} (Word 433)  \textbf{ NOSTALGIA} for their childhood.

\question \textbf{Conjecture}\\
\textit{ An inference based upon guesswork; a} \textbf{SUPPOSITION.}\\

$\cdot$ What caused the sudden extinction of the dinosaurs? Scientists have offered a number of \textbf{CONJECTURES} to explain why the Age of Dinosaurs came to an abrupt end. One popular \textbf{CONJECTURE} suggests that a giant meteor struck Mexico's Yucatan Peninsula, causing wide-spread fire storms, tidal waves, and the severe downpour of acid rain. An alternative \textbf{CONJECTURE} suggests that massive volcanic eruptions at the Deccan Flats in India caused climate changes that killed the dinosaurs. While both \textbf{CONJECTURES} are \textbf{PLAUSIBLE} (Word 38), scientists still lack a definitive explanation. 

\question \textbf{OBSOLETE, ARCHAIC, ANTIQUATED} \\ \textit{No longer in use; outmoded in design or style.}\\

$\cdot$ For many years Kodak was the \textbf{ICONIC} (idolized as an object of attention or devotion) leader in the photo industry. Many of its products became \textbf{ANTIQUATED} and, in the case of camera film, nearly \textbf{OBSOLETE}. Kodak's \textbf{MYOPIC} (shortsighted, lacking foresight) business model caused them to be late in entering the successor market—digital photography.

\question \textbf{AUSPICIOUS, PROPITIOUS}\\
\textit{Very  favorable.}\\

$\cdot$How long would you wait to marry your true love? The Mogul princes of India were required to wait until the emperor's astrologers felt that all of the planetary signs were \textbf{AUSPICIOUS}. For example, they required Crown Prince Shah Jahan and Mumtaz Mahal to postpone their wedding date for five years. During that time, the lovers were not allowed to see one another. The long-awaited wedding finally took place when all of the astrological signs were \textbf{ AUSPICIOUS}. The signs must have indeed been \textbf{PROPITIOUS} because the royal couple enjoyed 19 years of marital joy and happiness.
\question  \textbf{GAFFE}\\ 
\textit{A blunder; a faux pas; a clumsy social or diplomatic error}\\

$\cdot$ The 2012 Olympic Games provided their share of \textbf{GAFFES}. Just before the soccer events began, it was learned that the keys to Wembley Stadium had been lost, forcing officials to hastily change all the locks. It appears that the keys had not been stolen, just misplaced. Then the North Korean women's soccer team walked off the field at their opening match when organizers mistakenly introduced the players displaying South Korea's flag on the stadium screens. This was a serious faux pas: the two countries are still technically at war. Only after more than an hour's coaxing, \textbf{ABJECT} (humble) apologies, and the replacement of South Korea's largely white flag with images of North Korea's red banner did the offended North Korean women agree to take the field. Another embarrassing blunder occurred when the New Zealand Olympic Committee forgot to register the defending champion Valerie Adams for the shot put. The \textbf{GAFFE} was spotted before it was too late, and her name was added to the roster.
\question \textbf{IMPASSE}\\ 
\textit{A deadlock; stalemate; failure to reach an agreement}\\

$\cdot$ In The Hunger Games, the Gamemakers change the rules and announce that two tributes from the same district may win the competition together, so District 12 tributes Katniss and Peeta team up to defeat the others. When they are the only remaining tributes, the Gamemakers \textbf{RESCIND} (revoke) the previous rule change and say that only one of them can win in the deadly competition. In response, Katniss takes some poisonous berries from her pouch and shares them with Peeta; they intend to eat the berries together rather than fight each other. Katniss and Peeta are at an \textbf{IMPASSE} with the Gamemakers. They would rather die together than fight, and the Gamemakers want only one victor. Finally, the Gamemakers are \textbf{COERCED} (Word 273) into allow ing both victors because of Katniss and Peeta's suicide threat. They would rather have two winners than none.

\question \textbf{ANACHRONISM}\\ 
\textit{The false assignment of an event, person, scene, or language to a time when the event, person, scene, or word did not exist}\\

$\cdot$ Northern Renaissance artists often included \textbf{ANACHRONISMS} in their paintings. For example, Last Supper by the 15th century artist Dirk Bouts shows Christ and his disciples eating in a royal palace in what is today Belgium. While the \textbf{ANACHRONISM} in Bouts's painting is deliberate, the \textbf{ANACHRONISMS} in modern movies are unplanned blunders. For example, in the Civil War movie Glory, a digital watch is clearly visible on the wrist of a boy waving goodbye to the black soldiers of the 54th Massachusetts Regiment. And in the movie Gladiator, you can see a gas cylinder in the back of one of the overturned \say{Roman} chariots!

\begin{tcolorbox}
\begin{center}
\textbf{ KNOW YOUR ROOTS}\\
\textbf{GREEK ROOT: CHRONO } - \textit{time}.\\  
\end{center}  
\textbf{CHRONOLOGY} - \textit{the science of recording events by date}\\ 
\textbf{CHRONIC} - \textit{continuing for a long time.}\\ \textbf{SYNCHRONIC} - \textit{happening at the same time}\\ 
\textbf{SYNCHRONICITY} -\textit{phenomenon of events which coincide in time and appear meaningfully related but have no discoverable causal connection}\\ \textbf{SYNCHRONIZE} -  \textit{to cause to go at the same rate or occur at the same time (as a timepiece or a schedule)}\\ 
\textbf{CHRONICLE} -  \textit{a record of events in order of time}\\ 
\textbf{CHRONICLER} -  \textit{an historian, as a chronicler of event}
\end{tcolorbox}

\question \textbf{BELIE}\\ \textit{an historian, as a chronicler of events}\\

$\cdot$ In Catching Fire, the second installment of The Hunger Games \textbf{SAGA} (Word 236), Katniss and Peeta are forced to return to the arena for the Quarter Quell, a special 75th edition of the Hunger Games, in which they must compete against other previous victors of the Games. They form alliances with several of the other tributes, including Wiress, an \textbf{ECCENTRIC} (Word 157) woman from District 3 who rarely speaks in complete sentences. Her \textbf{UNCONVENTIONAL} (Word 7) and seemingly unbalanced behavior has earned her the nickname \say{Nuts.} However, her unusual behavior \textbf{BELIES} an extraordinary intelligence and intuition. She becomes a strong asset to the team, figures out crucial information concerning the arena's design, and helps her allies survive in the dangerous environment of the Games.

\question \textbf{MITIGATE, MOLLIFY, ASSUAGE, ALLEVIATE}\\ \textit{To relieve; to lessen; to ease.}\\

$\cdot$ Did you know that almost half of all Americans take at least one prescription pill every day? Americans use pills to \textbf{ALLEVIATE} the symptoms of everything from migraine headaches to acid indigestion. Stephen Douglas believed that the doctrine of popular sovereignty would \textbf{MITIGATE}, or lessen, the public's passions against the extension of slavery into the territories. But Douglas badly misjudged the public mood in the North. Instead of \textbf{MOLLIFYING} the public, popular sovereignty inflamed passions and helped propel the nation toward the Civil War.

\question \textbf{COVET}\\
\textit{To strongly desire; to crave}\\ \textbf{COVETOUS}\\ \textit{Grasping, greedy, eager to obtain something;} \textbf{AVARICIOUS} (Word 255)\\
 
$\cdot$ What do Lord Voldemort ( Harry Potter and the Deathly Hallows), The Wicked Witch of the West ( Wizard of Oz ), and Megatron ( Transformers) all have in common? All three villains are \textbf{COVETOUS} of something they desperately want but can’t have. Lord Voldemort \textbf{COVETS} the Elder Wand, the Wicked Witch of the West \textbf{COVETS} Dorothy's Ruby Slippers, and Megatron \textbf{COVETS} the All Spark.

\question \textbf{ANTITHESIS}\\ \textit{The direct or exact opposite; extreme contrast;} \textbf{ANTIPODE}\\ \textbf{ANTITHETICAL}\\ \textit{Exactly opposite;} \textbf{ANTIPODAL}\\
 
 $\cdot$ In her song \say{You Belong With Me,} Taylor Swift cannot \textbf{FATHOM} (understand) why a guy she likes continues to go out with a girl who is his complete \textbf{ANTITHESIS}. Their tastes in music and sense of humor are \textbf{ANTITHETICAL}. But Taylor recognizes that her rival is a cheer captain who \say{wears short skirts} while Taylor sits in the bleachers and  \say{wears t-shirts.} All Taylor can do is hope that the guy will have an \textbf{EPIPHANY} (Word 327) and realize that they belong together.

\question \textbf{PROTOTYPE}\\ 
\textit{An original model; an initial design}\\ 

$\cdot$ What do the Model T and The Bat in The Dark Knight Rises have in common? Although very different vehicles, both were originally designed to be \textbf{PROTOTYPES}. The Model T, invented by Henry Ford in 1908, served as the \textbf{PROTOTYPE} for the world's first affordable, mass-produced automobile. The Bat, created by Luciuslef tf Fox at Wayne Enterprises, was a \textbf{PROTOTYPE} for a flying tank military vehicle, but it helped Batman save Gotham from Bane and his men.

\question \textbf{ALOOF}\\
\textit{Detached; distant physically or emotionally; reserved; standing near but apart}\\

$\cdot$ In The Great Gatsby, Fitzgerald initially portrays Jay Gatsby as the \textbf{ALOOF} host of lavish parties given every week at his \textbf{ORNATE} (Word 397) mansion. Although he is courted by powerful men and beautiful women, Gatsby chooses to remain distant and \textbf{ALOOF}.\\

 In Homer's Iliad, many people accused Zeus of \say{wanting to give victory to the Trojans.} But Zeus chose to remain ALOOF: \say{He sat apart in his all-glorious majesty, looking down upon the Trojans, the ships of the Achaeans, the gleam of bronze, and alike upon the slayers and the slain.}
 
\question \textbf{TRITE, HACKNEYED, BANAL, PLATITUDINOUS, INSIPID}\\ 
\textit{Unoriginal; commonplace; overused; CLICHED}\\

$\cdot$ In The Catcher in the Rye, Holden Caulfield just can't help seeing most people as \say{phony}-his favourite word. When he goes to hear Ernie, the jazz piano player, he thinks of the playing as \textbf{BANAL}: so lacking in originality that it is almost boring. He sees straight through his headmaster’s \textbf{PLATITUDE} that \say{Life is a game,} understanding the message to be \textbf{TRITE}, unoriginal, and lacking freshness. Many people who read The Catcher in the Rye today think of Holden Caulfield’s very character as \textbf{HACKNEYED}, because he represents a character we have seen all too many times: the moody, \textbf{DISAFFECTED} (disconnected), disgruntled teenager. But back in 1951, when the novel was first published, Salinger's portrait of a young person was considered \textbf{SEARINGLY} (scorchingly) original. Paula Abdul, the former American Idol and X-Factor judge, was known for being nice and \textbf{AFFABLE} (Word 18), always saying something positive to the contestants. Although Paula was nice, her comments were \textbf{TRITE}, \textbf{BANAL}, and \textbf{HACKNEYED}. According to \textbf{PLATITUDINOUS} Paula, every singer was \say{great,} \say{beautiful,} and \say{amazing.} She encouraged each one with pleasant but \textbf{INSIPID} compliments like \say{You're authentic,} \say{America loves you,} and \say{Your journey of magic is just beginning.}

\question \textbf{ANTECEDENT}\\
\textit{A preceding event; a FORERUNNER; a PRECURSOR gnify.}\\  

$\cdot$ Many critics have noted that the 1995 Disney movie Pocahontas can be viewed as a thematic \textbf{ANTECEDENT} to the 2010 blockbuster Avatar.  In Pocahontas, \textbf{AVARICIOUS} (Word 255) English settlers search for gold. In Avatar, an \textbf{AVARICIOUS} company wants to mine unobtanium from the fictional planet Pandora. In both movies beautiful \textbf{INDIGENOUS} (Word 47) women rescue soldiers who find themselves drawn to the native peoples they originally intended to conquer. By helping Captain John Smith discover the New World's life and beauty, Pocahontas serves as an \textbf{ANTECEDENT} for Avatar’s Neytiri.

\begin{tcolorbox}
\begin{center}
\textbf{ KNOW YOUR ROOTS}\\
\textbf{GREEK ROOT: ANTE }- \textit{before.}\\ 
\end{center}  
 
\textbf{ANTEBELLUM} - \textit{before the Civil War}\\ \textbf{ANTEDILUVIAN} - \textit{before the Biblical flood, a hyperbolic word describing something extremely old.}\\ \textbf{ANTEDATE} - \textit{to precede in time.}\\
\textbf{ANTEROOM} - \textit{a waiting room outside a larger room.}\\ 
\textbf{ANTERIOR} - \textit{before in time and place.}
\end{tcolorbox}

\question \textbf{PLAUSIBLE}\\ \textit{Believable; credible}\\ 
\textbf{IMPLAUSIBLE}\\
\textit{Unbelievable; incredible}\\ 

$\cdot$ Let's play \textbf{PLAUSIBLE} or \textbf{IMPLAUSIBLE}: In the Bourne Ultimatum, Jason Bourne successfully breaks into Noah Vosen's heavily-guarded top- security office and steals an entire set of classified Blackbriar documents. \textbf{PLAUSIBLE} or \textbf{IMPLAUSIBLE}? \textbf{PLAUSIBLE}—because he is Jason Bourne! I n The Avengers, Iron Man, Captain America, Thor, The Hulk, Hawkeye, and Black Widow successfully save New York City from an extraterrestrial attack and a nuclear missile. \textbf{PLAUSIBLE} or \textbf{IMPLAUSIBLE}? \textbf{PLAUSIBLE}-because The Avengers all have special skills and powers that allow them to defeat their foes!

\question \textbf{PRUDENT}\\
\textit{Careful; cautious; sensible}\\

$\cdot$ In the Twilight \textbf{SAGA} (Word 236), Bella Swan is a high school student who meets and falls in love with Edward Cullen. However, Edward is not just another high school student. He is a 107-year-old vampire who stopped aging physically at 17. Edward understands that their relationship poses grave dangers to Bella. However, Bella and Edward love each other, so they decide to stay together despite the danger. Together, they must be PRUDENT in dealing with the dangers that they face, among them werewolves, vengeful vampires, and the \textbf{OMINOUS} (Word 197) Volturi

\question \textbf{AESTHETIC}\\
\textit{Relating to the nature of beauty, art, and taste; having a sense of what is beautiful, attractive, or pleasing.}\\

$\cdot$ Do you know why the Mona Lisa is considered one of the most beautiful paintings of all time? The answer lies in its use of the Golden Ratio, the naturally occurring ratio of height to width that is most \textbf{AESTHETICALLY} pleasing to humans. The Mona Lisa's face is composed entirely of Golden Ratio rectangles and thus adds to the overall \textbf{AESTHETIC} of the painting. However, the Golden Ratio is not limited to art. Examples can be found in ancient Greek architecture, Egyptian pyramids, biology, and even widescreen television screens! It is not \textbf{AESTHETICALLY} pleasing if a character introduced at the very end solves a novel or play's conflicts. Aristotle criticized Euripides' play Medea for having Medea saved at the end by a character not integral to the plot. To his mind, \textbf{AESTHETICALLY} this was not a satisfying conclusion.

\question \textbf{PARADOX}\\ 
\textit{A seemingly contradictory statement that nonetheless expresses a truth}\\

$\cdot$ One of the most famous literary first lines is that of Charles Dickens's A Tale of Two Cities : \say{It was the best of times, it was the worst of times.} How could such a contradiction be true? In the course of the book, this \textbf{PARADOXICAL} statement is shown to be valid. In Mary Shelley's novel Frankenstein, the creature encounters many \textbf{PARADOXES}. One is the simultaneous positive and negative characteristics of fire. It can warm him, protect him, light his way, and cook his food, but it can also burn and destroy. Similarly, the creature also comes to recognize the \textbf{PARADOXICAL} nature of man: driven by conflicting forces of selfishness and \textbf{ALTRUISM} (Word 20).
\question \textbf{ENIGMATIC, INSCRUTABLE}\\ \textit{Mysterious; puzzling; unfathomable; baffling}\\ 

$\cdot$What do Da Vinci's Mona Lisa, Fitzgerald's description of Jay Gatsby, and J.K. Rowling's portrayal of Snape have in common? All three figures are ENIGMATIC. The Mona Lisa's \textbf{ENIGMATIC} smile has puzzled art lovers for centuries. When The Great Gatsby opens, Jay Gatsby is an \textbf{ENIGMATIC} figu IN tre whose great wealth and extravagant parties spark endless gossip. And Snape's personality and loyalties remain \textbf{INSCRUTABLE} until the final chapters of Harry Potter and the Deathly Hallows. 
\question \textbf{ACQUIESCE}\\ \textit{To comply; agree; give in.}\\ 

$\cdot$ In Pirates of the Caribbean: The Curse of the Black Pearl, Elizabeth Swann and Captain Barbossa conduct negotiations that include long words. Elizabeth Swann: Captain Barbossa, I am here to negotiate the cessation of hostilities against Port Royal. Captain Barbossa: There be a lot of long words in there, Miss. We're naught but humble pirates. What is it that you want? Elizabeth Swann: I want you to leave and never come back. Captain Barbossa: I'm disinclined to \textbf{ACQUIESCE} to your request. Means no!  Although he is a \say{humble pirate,} Captain Barbossa can use long words as well as she can.

\question \textbf{NAÏVE, GULLIBLE}\\ 
\textit{simple; lacking worldly expertise; overly \textbf{CREDULOUS}; unsophisticated; immature; inexperienced; \textbf{INGENUOUS} (Word 428)}\\  

$\cdot$ Nemo, of Finding Nemo, is a young clown fish who thinks he is old enough to swim out in the open waters. Young, NAÏVE, and wanting to defy his overprotective father, he wanders too near a boat. Suddenly, a net surrounds him. He is taken aboard the boat and from there to Sydney, Australia, to live in a fish tank. His father Marlin, \textbf{DESPONDENT} (Word 176) at his loss, vows to find his son. Marlin succeeds and ultimately brings Nemo back home. By the end of the film, Nemo has learned the importance of obeying his father and of not being so \textbf{GULLIBLE}.

\question \textbf{AUTONOMY}\\ \textit{Independence; self-governance} \\ \textbf{AUTONOMOUS}\\ \textit{Acting independently, or having the freedom to do so; not controlled by others.}\\ 

$\cdot$ Fahrenheit 451, the classic novel by Ray Bradbury, imagines a \textbf{DYSTOPIA} (an imaginery society characterized by oppression and human misery) society in which a faceless government exerts huge control over its citizens. No books are allowed; instead, citizens watch endless television streams of \textbf{PROPAGANDA} (zealous advancement of a group's principles) from the government. Bradbury's novel suggests that people naturally den \textbf{AUTONOMY} in their own lives; if a faceless government tries to exert authority over them, they will tend to be \textbf{SUBVERSIVE} (tending to overthrow), and rebel against that authority. In the movie Men in Black, Agent Zed explains that MIB is an \textbf{AUTONOMOUS} organization that is \say{not a part of the system.} He goes on to say that MIB is \say{above the system, over it, beyond it; we are they, we are them, we are the Men in Black.} They are serious about their \textbf{AUTONOMY}.
\question \textbf{FUTILE}\\ \textit{Completely useless; doomed to failure; in vain.}\\

$\cdot$ The Deepwater Horizon oil spill released a \textbf{PRODIGIOUS} (huge, massive) flood of crude oil into, the Gulf of Mexico. BP engineers made repeated attempts to control or stop the spill. However, all of their initial efforts proved to be \textbf{FUTILE}. Although crews worked tirelessly to protect hundreds of miles of beaches, wetlands, and estuaries, local residents worried that these efforts would also prove to be \textbf{FUTILE}.
\question \textbf{INDIGENOUS, ENDEMIC}\\ \textit{Native to an area}\\ 

$\cdot$ Which of the following are Old World plants and animals, and which are New World plants and animals: potatoes, tomatoes, maize, sunflowers, cocoa beans, turkeys, and buffaloes? Surprisingly, all of these plants and animals are \textbf{INDIGENOUS} or \textbf{ENDEMIC} to the New World.
\begin{tcolorbox}
\begin{center}
\textbf{ KNOW YOUR ROOTS}
\end{center}  
\textbf{GREEK ROOT: DEM, DEMO} | the people.\\  \textbf{PANDEMIC} (Word 49)  of all the people, prevalent over a whole area.\\ 
\textbf{DEMOCRACY} - rule by the people, by the majority.\\ 
\textbf{DEMAGOGUE} (Word 111) a person who tries to stir up the people by appealing to emotion and prejudice in order to achieve selfish ends.\\ \textbf{DEMOGRAPHICS} (Word 433) - the science of vital statistics about populations (births, deaths, marriages, incomes, etc.)\\ 
\textbf{EPIDEMIC} - a rapid spread of a contagious disease or other negative condition.
\end{tcolorbox}
\question \textbf{UBIQUITOUS, PREVALENT}\\ \textit{Characterized by being everywhere; omnipresent; widespread}; \textbf{PERVASIVE}.\\ 

$\cdot$ What do cell phones, iPods, Starbucks The coffee shops, and McDonald's fast-food restaurants have in common? They are all \textbf{UBIQUITOUS}-we see them everywhere. Popular fashions are also \textbf{PERVASIVE}. For example, baggy knee-length shorts have completely replaced the once-\textbf{PREVALENT} short shorts of the 1970s. From high school b-ballers to WNBA and NBA superstars, long shorts are now \textbf{UBIQUITOUS}.
\question \textbf{PANDEMIC} -  \textit{An epidemic that is geographically widespread and affects a large proportion of the population.}\\ 

 $\cdot$ In the movie I Am Legend, a manmade virus known as KV triggers a global \textbf{PANDEMIC} that kills almost all of the human population on Earth. While there has never been a real \textbf{PANDEMIC} of this magnitude, virus strains and diseases have caused widespread deaths. In 1347 the Black Plague killed as many as one-third of the people in Europe. In the 16th century, Spanish conquistadores spread small pox and other diseases that \textbf{DECIMATED} (destroyed a great proportion of) the \textbf{INDIGENOUS} (Word 47) populations in Central America, the Caribbean, and Mexico. Our own times have not been immune to epidemics. The 1918 flu PANDEMIC killed 50 to 100 million people, and more recently we have had SARS, Asian Bird Flu, and Swine Flu PANDEMICS.\\ \textbf{PANDEMIC} can also be used as an adjective, meaning \textbf{PREVALENT} (Word 48) over a large area.
\question \textbf{FORTITUDE}\\  \textit{Strength of mind that allows one to endure pain or adversity with courage}.\\ 

$\cdot$ William Lloyd Garrison and Rosa Parks demonstrated great personal \textbf{FORTITUDE}. While most Americans accepted slavery, Garrison boldly demanded the immediate and unconditional emancipation of all slaves. Although initially ignored, Garrison \textbf{PERSEVERED} (refused to give up no matter the situation) and lived to see President Lincoln issue the Emancipation Proclamation. Rosa Parks also illustrates the principle that \textbf{FORTITUDE} is needed to achieve difficult goals. While most Americans accepted segregation, Rosa refused a bus driver's order to give up her seat to a white passenger. Her historic action helped \textbf{GALVANIZE} (Word 148) the Civil Rights Movement.
\clearpage
51.Upbraid, Reproach, Castigate-to scold, rebukeDirect Hits: Part 2
52.Diminutive-very small
53.Archaic-out of date. old fashoined
54.Exhort-a strongly encourge
55.Antipathy \& Animosity- strong dislike
56.Digress-showing great ditermination
57.Indulgent-overly tolerent
58.Polarize \& Divisive- to break into opposing fractions
59.Neboulus-vague, lacking a fully developed form
60.Analogy \& Analogous-a similarity or likeness
61.Fleeting \& Ephemeral-very brief, short lived-
62.Penchant \& Predilection-a prefference for something, an inclination-
63.Boorish \& Uncouth-vulgar , crude
64.Capricius \& Mercurial-ficle , Constantly shifting moods-
65.Indignant-outrate at something that injust-
66.Innuendo-a veiled refference-
67.Thwart \& Stymie-to stop, frustate
68.Adroit \& Deft \& Dexterous-skillful
69.Admonish-to earnestly caution-
70.Incontrovertible-indisputable, beyond doubt
71.Voracious \& Ravenous-a huge appetite, cannot be satisfied,insatiable-
72.Callous-insensitive-
73.Intrepid \& Undaunted-fearless ,courageous
74.Nonchalant-casual indifference-͢
75.Convoluted-twisted, intricate-̲
76.Itinerant-mobile, not sedantary
77.Poignant-touching, heartrending
78.Impetus-a stimulus or encourgement-
79.Bucolic,Rustic,Pastoral-charmingly rural
80.Equanimity-calmness,composure-
81.Panache\& Verve-dash and flamboyant-
82.Provocative-Provokes controversy
83.Placid \& Serene-very calm, quiet
84.Fortuitous-an accidental but fortune occurence-
85.Dispel-to drive away, scatter-
86.Amalgam-a mixure; combination of different elements
87.Viable \& Feasible-possible
88.Anguish-agonizing physical or mental pain
89.Intemperate-lacking restraint , excessive
90.Superficial-shallow; lacking depth-
91.Laud, Extol, Tout, Accalaim-praise, applaud- 
92.Dismissive-to reject, disregard
93.Disparage- belittle, slight
94.Pompous-pretentious, field with exceesive shelf importance-
95.Cryptic-mysterious, having a hidden meaning
96.Subtle-a gradual almost imperceptible change-
97.Disparity-an inequality, imbalance
98.Curtail- a cut short or reduce
99.Innocuous-harmless
100.Diatrible \& Tirade-a bitter denunciation- 
101. CHARLATAN-a fake; a fraud; a cheat
102. SKEPTIC - a doubter
103. RHETORICIAN- an eloquent writer or speaker
104. HEDONIST-seeker of pleasure
105. ASCETIC-a person who leads a life of self-denial
106. RACONTEUR-a person who excels in telling anecdotes
107. ICONOCLAST-someone who attacks cherished ideas and institutions
108. DILETTANTE - an amateur or dabbler
109. PARTISAN - a person with biased beliefs
110. MENTOR-a teacher; a guide; an advisor ACOLYfE-a devoted student
111. DEMAGOGUE - a speaker who appeals to emotions, fears or prejudices
112. AUTOMATON-a person who acts in a mechanica1 fashion; a mindless follower
113. RECLUSE - a person who leads a secluded, solitary life
114. BUNGLER-a clumsy or inept person
115. CLAIRVOYANT - a person who uses intuition to see into the future; a seer
116. PROGNOSTICATOR-a person who makes predictions based upon data
117. PUNDIT-a professional commentator
118. ZEALOT-a very enthusiastic person
119. NEOPHYTE, NOVICE and GREENHORN-a beginner
120. BENEFACTOR-a person who gives gifts BENEFICIARY-a person who receives benefits
121. DISSEMBLER and PREVARICATOR-a liar; deceiver
122. PROPONENT and ADVOCATE-a champion of a cause
123. PRODIGY - a young genius
124. ORACLE - a person who is a source of wise counsel and prophetic advice
125. MISANTHROPE-a person who bates humankind
126. INNOVATOR-a person who introduces something new
127. SYCOPHANT and OBSEQillOUS- a person who behaves in a servile manner; a toady
128. STOIC- a person who is impassive and emotionless
129. REPROBATE-a morally unprincipled person
130. RENEGADE-a disloyal person
131. DRACONIAN -very strict laws and rules
132. LACONIC, SUCCINCT, TERSE- very concise; brief
133. SPARTAN- plain; simple
134. HALCYON -idyllically calm and tranquil
135. SOPHISTRY-a deliberately misleading argument
136. CHIMERICAL - a fantastic scheme; unchecked imagination
137. OSTRACIZE-to deliberately exclude from a group
138. IMPECUNIOUS - poor; penniless; not affluent
139. NEFARIOUS- extremely wicked; vile
140. JOVIAL and JOCULAR-good-humored; cheerful
141. DIRGE -a funeral hymn; mournful music
142. MAUDLIN-excessively sentimental
143· QUIXOTIC-foolishly impractical
144· PANDEMONIUM - a wild uproar; tumult
145. MARTINET-a strict disciplinarian
146. FIASCO and DEBACLE-a complete failure
147. BOWDLERIZE- to remove or delete objectionable parts of a book
148. GALVANIZE - to electrify; stir into action
149. PICAYUNE-something of small value; petty; trifling
150. GERRYMANDER -to divide a district so as to give one side an advantageDirect Hits Vocabolury
151. MAVERICK-an independent person
152. JUGGERNAUT-an irresistible force
153. SERENDIPITY-an accidental but fortunate discovery
154. ZENITH and APEX - the highest point
155. NADIR-the lowest point
156. EXPUNGE~ EXCISE, EXPURGATE- delete; remove
157. ECCENTRIC-an odd, unconventional person
158. EXTRICATE-to get out of a difficult situation
159. EXEMPLARY-outstanding
160. ENUMERATE - to list; to tick off
161. ELUSIVE - out of reach; difficult to catch
162. EXORBITANT-unreasonably expensive
163. REPUDIATE, RECANT~ RENOUNCE -to take back; disavow
164. REDUNDANT-doing or saying something again and again
165. RELINQUISH - to give something back
166. RESILIENT - to bounce back
167. REAFFIRM - to assert again
168. RETICENT-to hold back one's thoughts and feelings
169. REBUFF - to repel or drive back; to reject
170. RENOVATE -to make new again
171. REJUVENATE - to make young again
172. RESURGENT-to rise again; surge back
173. DELETERIOUS -harmful; injurious
174. DECRY-to express strong disapproval
175. DESPONDENT and FORLORN- feeling downcast; dejected
176. DENOUNCE-to speak against
177. DEMISE-the final ending of something; the downfall
178. DEBUNK-to put down by exposing false claims
179. DERIDE - to put down with contemptuous jeering
180. DEVOID -completely lacking in something
181. IMPECCABLE -faultless; perfect
182. IMPLACABLE-not capable of being appeased
183. INEXORABLE -relentless, unstoppable
184. INCOHERENT-lacking organization or logic
185. INSURMOUNfABLE - not capable ofbeing overcome
186. IRREVERENT - lacking proper respect; disrespectful
187. CIRCUMSPECT and PRUDENT-cautious; careful
188. CIRCUITOUS - circular and therefore indirect
189. CIRCUMVENT- to avoid by artful maneuvering
190. CIRCUMSCRJBE - to narrowly restrict; to limit action; to draw a line around
191. MAGNANIMOUS -generous and forgiving
192. ERRONEOUS -filled with errors; wrong
193. MOMENTOUS -of great importance
194. MELLIFLUOUS -smooth and sweet flowing
195. OMINOUS -threatening and menacing
196. ACRIMONIOUS -great bitterness
197. COPIOUS-abundant; a great amount
198. ABSTEMIOUS-moderate in eating and drinking
199. MALODOROUS -foul smelling
200.TEDIOUS- boring and tiresome
\end{questions}
\end{document}