\documentclass{book}
\usepackage[utf8]{inputenc}
\usepackage{geometry,dirtytalk,amsmath}
\usepackage{tikz}
\usepackage{ tcolorbox,array}
\usetikzlibrary{scopes}
\usepackage{draftwatermark}
% \SetWatermarkLightness{0.8}
\SetWatermarkScale{.5}
\usepackage{lipsum}
\SetWatermarkText{Direct Hits}

\begin{document}
\chapter{Core Vocabulary (1-50)}
\say{I constantly see people rise in life who are not the smartest, sometimes not even the most diligent, but they are learning machines. They go to bed every night a little wiser than they were when they got up and boy does that help, particularly when you have a long run ahead of you.}

\hfill - \textbf{Charlie Munger}

The English language contains just over one million words-the most of any language in human history. If each of these words had an equal chance of being used on the SAT, studying for the test would be a truly impossible task. Fortunately, the pool of words used by Educational Testing Service (ETS) test writers is actually relatively small. Questions on the test are ranked by level of difficulty from 1 to 5, with 5 being the most difficult. In general, level 3 and 4 questions are missed by over half of the test-takers. These crucial mid-level words, the level 3 and 4 words, form the core \textbf{LEXICON} or special vocabulary you need to know to score well on the Critical Reading portion of the SAT. After a careful analysis of recent tests, we have identified 100 Core Vocabulary Words. The first 50 of these words are in Chapter 1, and the second 50 are in Chapter 2. The division is arbitrary. Each word is a high-frequency word that you absolutely must know
\begin{enumerate}
\item  \textbf{AMBIVALENT}\\
\textit{Contradictory, having mixed feelings.}\\

$\bullet$ In the TV show Glee, Finn Hudson is the star
quarterback of his high school football team. Finn is
also a talented singer who is \textbf{AMBIVALENT} about
joining the school Glee Club. Although Finn doesn't
want to alienate his teammates, he enjoys singing and
wants to follow his dream of becoming an entertainer.\\

$\bullet$ In the movie The Notebook, Allie has to choose
between Noah and Lon. She is emotionally torn by her
\textbf{AMBIVALENT} feelings as she tells Noah, \say{There is no
easy way, no matter what I do, somebody gets hurt}.
She later reiterates her \textbf{AMBIVALENT} feelings when
she tells Lon \say{When I'm with Noah I feel like one
person, and when I'm with you, 1 feel like someone
totally different}.

\begin{tcolorbox}
\begin{center}
\textbf{ KNOW YOUR ROOTS}\\
\line(1,0){250}\\
\textbf{LATIN PREFIX: AMBI }$ - $ \textit{Both}.\\  
\end{center}  
\textbf{AMBIDEXTROUS} $ - $ \textit{able to use both hands with equal ease, skillful, versatile}\\
\textbf{AMBIGUOUS} (Word 21) $ - $ \textit{having two or more possible meanings, doubtful, dubious,} \textbf{ EQUIVOCAL} (Word 210)\\
\textbf{AMBIVALENT} $ - $ \textit{ being simultaneously of two mind}
\end{tcolorbox}

\item  \textbf{ANOMALY}\\ \textit{Deviation from the norm, something that is}
\textbf{ATYPICAL}.\\

$\bullet$ In the Harry Potter series, Vernon Dursley prides
himself on being \say{perfectly normal, thank you very
much}. An \textbf{ANOMALY} is the last thing Dursley wants
in his life. In the opening chapter of Harry Potter and
the Sorcerer's Stone, however, Dursley notices several
strange \textbf{ANOMALIES} or what he calls \say{funny stuff}.
For example, be spots a cat that appears to be reading
a map. He also notices a number of odd people who
are dressed in colorful robes. And that is not all: flocks
of owls can be seen flying during the daytime.

\item \textbf{SARCASTIC, SARDONIC, SNIDE}\\ \textit{Mocking, derisive, taunting, and stinging.}\\

$\bullet$ Winston Churchill was famous for his \textbf{SARCASTIC}
and \textbf{SARDONIC} comments. Here are two well-known
examples:\\
Bessie Braddock: Sir, you are a drunk.\\
Churchill: Madame, you are ugly. In the morning I shall be sober, and you will still be ugly.\\
Nancy Astor·: Sir, if you were my husband I would give you poison.\\
Churchill: If I were your husband I would take it.\\
In the movie Avatar, Dr. Grace Augustine tells Jake,
\say{Just relax and let your mind go blank. That shouldn't
be too hard for you}. This \textbf{SNIDE} remark expresses
Grace's initial contempt for Jake.

\item \textbf{DEARTH and PAUCITY}\\ \textit{A scarcity or shortage of something.}\\

$\bullet$ A study by USA Today revealed that college football's top level teams have a \textbf{DEARTH} of minority coaches. Just 10 of the 120 Football Bowl Subdivision
universities have minority head coaches. There is also
a \textbf{PAUCITY} of minority assistant coaches. In contrast,
about one-fourth of the 32 NFL teams have minority
head coaches. Critics encourage the colleges to
\textbf{EMULATE} (imitate) the NFL's policy of requiring
teams to interview minority candidates for head
coaching opportunities.

\item \textbf{PRATTLE}\\ \textit{To speak in a foolish manner, to babble incessantly.}\\

$\bullet$ In the movie Office Space, Milton continuously
\textbf{PRATTLES} to himself about how he is abused by
management and bow his co-workers repeatedly
borrow his stapler. Here is an example of Milton
\textbf{PRATTLING} on and on:\\
\say{I don't care if they Lay me off either, because I
told, I told Bill that if they move my desk one
more time, then, then, I'm, I'm quitting, I'm
going to quit. And, I told Don too, because
they've moved my desk four times already this
year, and I used to be over by the window, and
I could see the squirrels, and they were
married, but then, they switched from the
Swingline to the Boston stapler, but I kept my
Swingline stapler because it didn't bind up as
much, and I kept the staplers for the Swingline
stapler and it's not okay, because if they take
my stapler then I'll set the building on fire $\cdots$}

\item \textbf{WRY} \\
\textit{Dry, humorous with a clever twist and a touch of irony;} \textbf{DROLL}\\

$\bullet$ Casel Di Principe is normally a sleepy town northwest of Naples, Italy. On a quiet Sunday a 6o-year-old man sat at a table playing cards -with his friends. Two
gunmen suddenly shattered the peace when they
drove down the town's main street and fired a barrage
of bullets, killing the old man. The assassination of an
important crime boss caused authorities to launch a
full-scale investigation. However, the witnesses refused to answer questions, claiming that they didn't see anything. Their collective silence prompted a \textbf{WRY} comment from one police officer that \say{the victim must
have been playing solitaire}.\\

\begin{tcolorbox}
\begin{center}
\textbf{Tip for Direct Hit}\\
\line(1,0){250}\\
\end{center}
A WRY sense of humor is very different from a
\textbf{JOCULAR} sense of humor. A \textbf{WRY} joke appeals
to your intellect and often produces a knowing
smile. In contrast, a \textbf{JOCULAR} joke appeals to
your funny bone and produces a belly laugh.
\end{tcolorbox}

\item \textbf{UNCONVENTIONAL} and
\textbf{UNORTHODOX}\\ \textit{Not ordinary or typical; characterized by
avoiding customary conventions and behaviors.}\\

$\bullet$ Both Katy Perry and Lady Gaga are known for their
catchy hits and bold, \textbf{UNCONVENTIONAL}
wardrobes. Katy Perry's colorful, \textbf{UNORTHODOX}
costumes include a funky playing cards dress, a
watermelon dress, and a cute pink panda outfit.\\

Lady Gaga is also known for wearing
\textbf{UNCONVENTIONAL} and even \textbf{OUTLANDISH}
(bizarre, outrageous) stage outfits. Gaga is usually
attired in her trademark platinum blonde hair, funky
sunglasses, and revealing leotards. However, she is
always unpredictable. Gaga opened one concert
wearing clear plastic champagne bubbles and then
changed into a sculpted prom dress.

\item \textbf{PAINSTAKING} and \textbf{METICULOUS}\\ \textit{Extremely careful; very} \textbf{EXACTING}\\

$\bullet$ The new Harry Potter theme park is a \textbf{METICULOUS}
recreation of Hogwarts castle and nearby Hogsmeade
village. The park's designers spared no expense to
\textbf{PAINSTAKINGLY} recreate such iconic rooms as
Dumbledore's office and the Defense Against the Dark
Arts classroom. \textbf{ENTHRALLED} (fascinated) visitors
can sample butterbeer and even purchase a wand at
Ollivander's Wand Shop.

\item \textbf{AUDACIOUS}\\
\textit{Fearlessly, often recklessly daring; very bold.}\\

$\bullet$ What do American General George Washington and
Japanese Admiral Isoroku Yamamoto have in
common? Both launched \textbf{AUDACIOUS} surprise
attacks on unsuspecting adversaries. On Christmas
Day, 1776, Washington ordered the Colonial Army to
cross the Delaware and attack the British and Hessian
forces at Trenton. Washington's \textbf{AUDACIOUS} plan
shocked the British and restored American morale.
On December 7, 1941, Yamamoto ordered the
Japanese First Air Fleet to launch a surprise attack on
the American Pacific Fleet based at Pearl Harbor.
Although Japan's \textbf{AUDACIOUS} sneak attack
temporarily \textbf{HOBBLED} (hampered) the U.S. fleet, it
aroused the now unified country to demand revenge.

\item  \textbf{INDIFFERENT and APATHETIC}\\ \textit{Marked by a lack of interest or concern.}\\

$\bullet$ In the movie Ferris Bueller's Day Off, the economic
teacher Ben Stein delivers a \textbf{SOPORIFIC} (sleep inducing) lecture on tariffs and the Great Depression.
Stein's bored and \textbf{INDIFFERENT} students ignore his
monotone lecture. Hoping for some sign of interest,
Stein tries asking questions, but his efforts are
\textbf{FUTILE} (Word 46). Some students are so
\textbf{APATHETIC} they fall asleep.

\item \textbf{DIFFIDENT}\\
\textit{Lacking self-confidence; self-effacing; NOT assertive.}\\

$\bullet$ Even Lady Gaga wakes up and feels like an insecure and \textbf{DIFFIDENT} 24-year-old girl. But Gaga then tells herself, \say{You're Lady Gaga, you get up and walk the walk today.} As you study for the SAT, be like Lady
Gaga. Don't \textbf{SUCCUMB} (give in) to feelings of
\textbf{DIFFIDENCE}. Study your Direct Hits vocabulary and
be confident.

\item \textbf{PRAGMATIC}\\ \textit{Practical; sensible; NOT idealistic or romantic.}\\

$\bullet$ What do the 16th century French king Henry N and
the 20th century American president Franklin Delano Roosevelt have in common? Both leaders made \textbf{PRAGMATIC} decisions that helped resolve a crisis. Henry IV was the newly crowned Protestant king in a country dominated by Catholics. For the sake of his war-weary country, Henry IV \textbf{PRAGMATICALLY} chose to become a Catholic, saying, \say{Paris is worth a Mass}.\\
FDR was a newly-elected president in a country facing
the worst economic crisis in its history. For the sake of
his country, Roosevelt \textbf{PRAGMATICALLY} chose to replace traditional laissez-faire economic policies with
\say{bold, persistent experimentation}. FDR \textbf{PRAGMATICALLY} explained, \say{It is common sense to take a method and try it; if it fails, admit it frankly and try
another. But above all, try something}.

\item  \textbf{EVOCATION}\\
\textit{An imaginative re-creation.}\\

$ \bullet $ What do the treasures of Pharaoh Tutankhamen,
Taylor Swift's music video \say{Love Story}, and the
movie Titanic all have in common? They are all
powerful \textbf{EVOCATIONS}. The treasures of Pharaoh
Tutankhamen are \textbf{EVOCATIONS} of the power and
splendor of Ancient Egypt. Taylor Swift's \say{Love Stacy}
\textbf{EVOKES} a time when beautiful princesses lived in
romantic castles and fell in love with handsome
princes. And the movie Titanic is a remarkable
\textbf{EVOCATION} of what it was like to be a passenger on
the great but doomed ship.

\begin{tcolorbox}
\begin{center}
\textbf{ KNOW YOUR ROOTS}\\
\line(1,0){250}\\
\textbf{LATIN ROOT: VOC, VOK} $ - $ \textit{Call}.\\  
\end{center}  
\textbf{VOCAL} $ - $ \textit{related to the voice, speaking}\\
\textbf{VOCATION} $-$ \textit{your calling, your profession, often used for a religious career}\\
\textbf{AVOCATION} $-$  \textit{a second calling, a hobby}\\ 
\textbf{EVOKE} $-$  \textit{to call forth, especially from the past}\\
\textbf{REVOKE} $-$  \textit{to call back, to rescind, to repeal}\\ 
\textbf{INVOKE} $-$ \textit{to call upon. Epic poems often begin with an Invocation of the Muse, or goddess of artistic inspiration}\\
\textbf{PROVOKE} $-$  \textit{to call forth} (see Word 82)\\ 
\textbf{CONVOCATION} $-$ \textit{a calling together, a gathering}\\
\textbf{VOCIFEROUS} $-$ \textbf{making an outcry, clamorous}\\
\textbf{EQUIVOCATE} $-$  \textit{to use} \textbf{AMBIGUOUS} (Word 21) \textit{expressions, to mislead} \textbf{IRREVOCABLE} \textit{incapable of being recalled or altered.}
\end{tcolorbox}

\item \textbf{PRESUMPTUOUS}\\
\textit{Overbearing; impertinently bold; characterized by brashly overstepping one's place.}\\

$ \bullet $ In the movie 300, Queen Gorgo boldly told the
Persian envoy, \say{Do not be coy or stupid Persian. You
can afford neither in Sparta.} Queen Gorgo's willingness to speak out astonished the Persian envoy.
Shocked by the Spartan Queen's \textbf{PRESUMPTUOUS}
statement, the envoy questioned, \say{What makes this
woman think she can speak among men?}

\item \textbf{RECALCITRANT}\\
\textit{Stubborn resistance to and defiance of authority or guidance;} \textbf{OBSTINATE; OBDURATE.}\\

$ \bullet $ What do Hester Prynne (The Scarlet Letter) and the
singer Amy Winehouse have in common? Both are
\textbf{RECALCITRANT}. In The Scarlet Letter, the Reverend
Wilson demanded that Hester reveal the name of the
father of her child. But Hester was \textbf{RECALCITRANT}.
Despite \say{the heavy weight of a thousand eyes, all
fastened upon her,} Hester stubbornly refused to
name the father, defiantly declaring, \say{Never...! will
not speak!} In her song \say{Rehab,} Amy Winehouse is
also defiantly \textbf{RECALCITRANT}. Her friends and
family all beg her to go to rehab, but Amy is
\textbf{OBDURATE} and defiantly declares, \say{No, no, no.}

\item \textbf{BOON}\\ \textit{A timely benefit; blessing.}\\ \textbf{Bane}\\
\textit{A source of harm and ruin.} \\

$ \bullet $ Fifty Cent was shot nine times and lived! Was the
shooting a \textbf{BANE} or a \textbf{BOON} for his career? At first it
was a \textbf{BANE} because the pain was excruciating and
Fiddy bad to spend weeks in a hospital recuperating.
But the shooting turned out to be a \textbf{BOON} for his
career because it \textbf{BOLSTERED} (reinforced) Fiddy's
\say{street cred} and attracted lots of publicity.

\item \textbf{CLANDESTINE} and \textbf{SURREPTITIOUS }\\
\textit{Secret; covert; not open;} \textbf{NOT ABOVEBOARD}.\\

$ \bullet $ What do the Men in Black (Men In Black),
Dumbledore's Army (Harry Potter and the Order of the Phoenix, and Sector Seven (Transformers) all
have in common? They are all \textbf{CLANDESTINE} groups
that conduct \textbf{SURREPTITIOUS} activities. The Men in Black \textbf{SURREPTITIOUSLY} regulate alien life forms on Earth. Dumbledore's Army teaches Hogwarts students how to defend themselves agajnst the Dark Arts.
And Sector Seven guards the mysterious All Spark and
keeps the body of Megatron permanently frozen.
\item \textbf{AFFABLE, AMIABLE, GENIAL,
GREGARIOUS}\\ \textit{All mean agreeable; marked by a pleasing
personality; warm and friendly.}\\

$ \bullet $ President Reagan was renowned for his \textbf{AFFABLE}
grace and \textbf{GENIAL} good humor. On March 6, 1981, a
deranged gunman shot the president as he was
leaving a Washington hotel. The injured but always
\textbf{AMIABLE} president looked up at his doctors and
nurses and said, \say{I hope you're all Republicans.} The
first words be uttered upon regaining consciousness
were to a nurse who happened to be holcling the
president's hand. \say{Does Nancy know about us?} the
president joked.

\begin{tcolorbox}
If you think you have heard the word \textbf{AMIABLE} before, you are probably right. The English word
\textbf{AMIABLE} contains the Latin root ami meaning
friend. You may have beard this root in the
French word ami and the Spanish word amigo.
\begin{center}
\textbf{ KNOW YOUR ROOTS}\\
\line(1,0){250}\\
\textbf{LATIN PREFIX: AMI} $-$ \textit{Friend}.\\  
\end{center}  
\textbf{AMITY} $-$ \textit{Friendship, harmony}\\
\textbf{AMICABLE} $-$ \textit{peaceable, harmonious}
\end{tcolorbox}

\item \textbf{AUSTERE}\\
\textit{Having no adornment or ornamentation; bare;
not} \textbf{ORNATE} (Word 363).\\ \textbf{AUSTERITY}\\ \textit{The trait of great self$-$denial; economy.}\\

$ \bullet $ Ancient Greek architects often used Doric columns to construct temples. For example, the Parthenon's \textbf{AUSTERE} columns conveyed strength and simplicity because they lacked ornamentation.\\

Although modern Greeks admire the \textbf{AUSTERE}
columns built by their ancestors, they vigorously
oppose new \textbf{AUSTERITY} measures that raise taxes
and cut social welfare programs. These \textbf{AUSTERITY}
measures have provoked massive protests.

\item \textbf{ALTRUISTIC}\\\textit{Unselfish concern for the welfare of others.}\\

$ \bullet $ Eleven$-$year$-$old Olivia Bouler sobbed uncontrollably
when she first saw pictures of oil$-$coated birds dying
along the Gulf coast. Rather than continue to cry,
Olivia wrote a Jetter to the Audubon Society offering
to help. Olivia volunteered to draw and sell watercolor
paintings o fbirds and give the profits to the Audubon
Society. Olivia's \textbf{ALTRUISTIC} campaign has helped
raise over \$ 160,000 to save Gulfbirds.

\item \textbf{AMBIGUOUS}\\
\textit{Unclear; uncertain; open to more than one interpretation; not definitive.}\\

$ \bullet $ Tbe final scene of the movie Inception is deliberately \textbf{AMBIGUOUS}. Leo DiCaprio's character, Dom Cobb,is \textbf{ELATED} (very happy) because he has found his
children and completed the seemingly impossible job
he was hired to do. But is all this real or is Dom
entrapped in yet another dream? Dom uses a metal
top to enable him to determine what is real and what
isn't. At the end of the film, Dom spins the top. What
will happen next? If the top keeps spinning, Dom is
dreaming. If it falls, things are real. We don't know
what happens because the ending is \textbf{AMBIGUOUS}.

\item  \textbf{UPBRAID, REPROACH, CASTIGATE}\\ \textit{To express disapproval; to scold; to rebuke; to }\textbf{CENSURE.}\\

$\bullet$ In this classic scene from Billy Madison, Ms. Vaughn \textbf{UPBRAIDS} Billy for making fun of a third grade student who is having trouble reading: \\Third Grader: Wa$-$wa-wa-once th-th-th-th-there wa-wa-wa-was a-a-a-a g-g-girl \\Billy Madison: Kid can't even read. \\Ernie: Cut it out, dude, you're gonna get us in trouble. \\Billy Madison: T-T-T-Today Junior! \\Billy Madison: OW! You're tearing my ear off! \\Veronica Vaughn : Making fun of a little kid for trying to read. Are you psycho? Do you not have a soul? You keep your mouth shut for the next two weeks or I'm going to fail you. End of story.

\item  \textbf{NOSTALGIA}\\
\textit{A \textbf{WISTFUL} (Word 206) sentimental longing for a place or time in the past.}\\

$\bullet$ A lifelong fan of The Muppets, Jason Segel was \textbf{NOSTALGIC} for his childhood, and he decided to  \textbf{REJUVENATE} (Word 171) the franchise by writing a new movie for them. Segel said, \say{We set out to make a Muppet movie that harkened back to the late-'70s, early-'80s Muppets that we grew up with.} It's been over a decade since The Muppets starred in a theatrical movie, and, likewise, in The Muppets, it's been a while since Kermit and his friends have performed as a group. As the audience revisits their childhood icons during this \textbf{NOSTALGIC} film, The Muppets, too, take a \textbf{WISTFUL} (Word 206) walk down memory lane. The Muppets decide to get their group together again for one last show, but they discover that they aren't popular anymore. They have become \textbf{ANTIQUATED} (Word 25); one character tells them, \say{You're \textbf{RELICS} (surviving objects from the past).} By incorporating clever humor and  \textbf{WISTFUL} (Word 206) references to Muppet movies of the past, The Muppets introduces a new generation to the \textbf{WHIMSICAL} (Word 219) world of Kermit and his friends while also catering to an older \textbf{DEMOGRAPHIC's} (Word 433)  \textbf{ NOSTALGIA} for their childhood.

\item \textbf{CONJECTURE}\\
\textit{ An inference based upon guesswork; a} \textbf{SUPPOSITION.}\\

$\bullet$ What caused the sudden extinction of the dinosaurs? Scientists have offered a number of \textbf{CONJECTURES} to explain why the Age of Dinosaurs came to an abrupt end. One popular \textbf{CONJECTURE} suggests that a giant meteor struck Mexico's Yucatan Peninsula, causing wide-spread fire storms, tidal waves, and the severe downpour of acid rain. An alternative \textbf{CONJECTURE} suggests that massive volcanic eruptions at the Deccan Flats in India caused climate changes that killed the dinosaurs. While both \textbf{CONJECTURES} are \textbf{PLAUSIBLE} (Word 38), scientists still lack a definitive explanation. 

\item \textbf{OBSOLETE, ARCHAIC, ANTIQUATED} \\ \textit{No longer in use; outmoded in design or style.}\\

$\bullet$ For many years Kodak was the \textbf{ICONIC} (idolized as an object of attention or devotion) leader in the photo industry. Many of its products became \textbf{ANTIQUATED} and, in the case of camera film, nearly \textbf{OBSOLETE}. Kodak's \textbf{MYOPIC} (shortsighted, lacking foresight) business model caused them to be late in entering the successor market—digital photography.

\item \textbf{AUSPICIOUS, PROPITIOUS}\\
\textit{Very  favorable.}\\

$\bullet$How long would you wait to marry your true love? The Mogul princes of India were required to wait until the emperor's astrologers felt that all of the planetary signs were \textbf{AUSPICIOUS}. For example, they required Crown Prince Shah Jahan and Mumtaz Mahal to postpone their wedding date for five years. During that time, the lovers were not allowed to see one another. The long-awaited wedding finally took place when all of the astrological signs were \textbf{ AUSPICIOUS}. The signs must have indeed been \textbf{PROPITIOUS} because the royal couple enjoyed 19 years of marital joy and happiness.
\item  \textbf{GAFFE}\\ 
\textit{A blunder; a faux pas; a clumsy social or diplomatic error}\\

$\bullet$ The 2012 Olympic Games provided their share of \textbf{GAFFES}. Just before the soccer events began, it was learned that the keys to Wembley Stadium had been lost, forcing officials to hastily change all the locks. It appears that the keys had not been stolen, just misplaced. Then the North Korean women's soccer team walked off the field at their opening match when organizers mistakenly introduced the players displaying South Korea's flag on the stadium screens. This was a serious faux pas: the two countries are still technically at war. Only after more than an hour's coaxing, \textbf{ABJECT} (humble) apologies, and the replacement of South Korea's largely white flag with images of North Korea's red banner did the offended North Korean women agree to take the field. Another embarrassing blunder occurred when the New Zealand Olympic Committee forgot to register the defending champion Valerie Adams for the shot put. The \textbf{GAFFE} was spotted before it was too late, and her name was added to the roster.
\item \textbf{IMPASSE}\\ 
\textit{A deadlock; stalemate; failure to reach an agreement}\\

$\bullet$ In The Hunger Games, the Gamemakers change the rules and announce that two tributes from the same district may win the competition together, so District 12 tributes Katniss and Peeta team up to defeat the others. When they are the only remaining tributes, the Gamemakers \textbf{RESCIND} (revoke) the previous rule change and say that only one of them can win in the deadly competition. In response, Katniss takes some poisonous berries from her pouch and shares them with Peeta; they intend to eat the berries together rather than fight each other. Katniss and Peeta are at an \textbf{IMPASSE} with the Gamemakers. They would rather die together than fight, and the Gamemakers want only one victor. Finally, the Gamemakers are \textbf{COERCED} (Word 273) into allow ing both victors because of Katniss and Peeta's suicide threat. They would rather have two winners than none.

\item \textbf{ANACHRONISM}\\ 
\textit{The false assignment of an event, person, scene, or language to a time when the event, person, scene, or word did not exist}\\

$\bullet$ Northern Renaissance artists often included \textbf{ANACHRONISMS} in their paintings. For example, Last Supper by the 15th century artist Dirk Bouts shows Christ and his disciples eating in a royal palace in what is today Belgium. While the \textbf{ANACHRONISM} in Bouts's painting is deliberate, the \textbf{ANACHRONISMS} in modern movies are unplanned blunders. For example, in the Civil War movie Glory, a digital watch is clearly visible on the wrist of a boy waving goodbye to the black soldiers of the 54th Massachusetts Regiment. And in the movie Gladiator, you can see a gas cylinder in the back of one of the overturned \say{Roman} chariots!

\begin{tcolorbox}
\begin{center}
\textbf{ KNOW YOUR ROOTS}\\
\line(1,0){250}\\
\textbf{GREEK ROOT: CHRONO } $ -$ \textit{time}.\\  
\end{center}  
\textbf{CHRONOLOGY} $-$ \textit{the science of recording events by date}\\ 
\textbf{CHRONIC} $-$ \textit{continuing for a long time.}\\ \textbf{SYNCHRONIC} $-$ \textit{happening at the same time}\\ 
\textbf{SYNCHRONICITY} $-$\textit{phenomenon of events which coincide in time and appear meaningfully related but have no discoverable causal connection}\\ \textbf{SYNCHRONIZE} $-$  \textit{to cause to go at the same rate or occur at the same time (as a timepiece or a schedule)}\\ 
\textbf{CHRONICLE} $-$  \textit{a record of events in order of time}\\ 
\textbf{CHRONICLER} $-$  \textit{an historian, as a chronicler of event}
\end{tcolorbox}

\item \textbf{BELIE}\\ \textit{an historian, as a chronicler of events}\\

$\bullet$ In Catching Fire, the second installment of The Hunger Games \textbf{SAGA} (Word 236), Katniss and Peeta are forced to return to the arena for the Quarter Quell, a special 75th edition of the Hunger Games, in which they must compete against other previous victors of the Games. They form alliances with several of the other tributes, including Wiress, an \textbf{ECCENTRIC} (Word 157) woman from District 3 who rarely speaks in complete sentences. Her \textbf{UNCONVENTIONAL} (Word 7) and seemingly unbalanced behavior has earned her the nickname \say{Nuts.} However, her unusual behavior \textbf{BELIES} an extraordinary intelligence and intuition. She becomes a strong asset to the team, figures out crucial information concerning the arena's design, and helps her allies survive in the dangerous environment of the Games.

\item \textbf{MITIGATE, MOLLIFY, ASSUAGE, ALLEVIATE}\\ \textit{To relieve; to lessen; to ease.}\\

$\bullet$ Did you know that almost half of all Americans take at least one prescription pill every day? Americans use pills to \textbf{ALLEVIATE} the symptoms of everything from migraine headaches to acid indigestion. Stephen Douglas believed that the doctrine of popular sovereignty would \textbf{MITIGATE}, or lessen, the public's passions against the extension of slavery into the territories. But Douglas badly misjudged the public mood in the North. Instead of \textbf{MOLLIFYING} the public, popular sovereignty inflamed passions and helped propel the nation toward the Civil War.

\item \textbf{COVET}\\
\textit{To strongly desire; to crave}\\ \textbf{COVETOUS}\\ \textit{Grasping, greedy, eager to obtain something;} \textbf{AVARICIOUS} (Word 255)\\
 
$\bullet$ What do Lord Voldemort ( Harry Potter and the Deathly Hallows), The Wicked Witch of the West ( Wizard of Oz ), and Megatron ( Transformers) all have in common? All three villains are \textbf{COVETOUS} of something they desperately want but can’t have. Lord Voldemort \textbf{COVETS} the Elder Wand, the Wicked Witch of the West \textbf{COVETS} Dorothy's Ruby Slippers, and Megatron \textbf{COVETS} the All Spark.

\item \textbf{ANTITHESIS}\\ \textit{The direct or exact opposite; extreme contrast;} \textbf{ANTIPODE}\\ \textbf{ANTITHETICAL}\\ \textit{Exactly opposite;} \textbf{ANTIPODAL}\\
 
 $\bullet$ In her song \say{You Belong With Me,} Taylor Swift cannot \textbf{FATHOM} (understand) why a guy she likes continues to go out with a girl who is his complete \textbf{ANTITHESIS}. Their tastes in music and sense of humor are \textbf{ANTITHETICAL}. But Taylor recognizes that her rival is a cheer captain who \say{wears short skirts} while Taylor sits in the bleachers and  \say{wears t-shirts.} All Taylor can do is hope that the guy will have an \textbf{EPIPHANY} (Word 327) and realize that they belong together.

\item \textbf{PROTOTYPE}\\ 
\textit{An original model; an initial design}\\ 

$\bullet$ What do the Model T and The Bat in The Dark Knight Rises have in common? Although very different vehicles, both were originally designed to be \textbf{PROTOTYPES}. The Model T, invented by Henry Ford in 1908, served as the \textbf{PROTOTYPE} for the world's first affordable, mass-produced automobile. The Bat, created by Luciuslef tf Fox at Wayne Enterprises, was a \textbf{PROTOTYPE} for a flying tank military vehicle, but it helped Batman save Gotham from Bane and his men.

\item \textbf{ALOOF}\\
\textit{Detached; distant physically or emotionally; reserved; standing near but apart}\\

$\bullet$ In The Great Gatsby, Fitzgerald initially portrays Jay Gatsby as the \textbf{ALOOF} host of lavish parties given every week at his \textbf{ORNATE} (Word 397) mansion. Although he is courted by powerful men and beautiful women, Gatsby chooses to remain distant and \textbf{ALOOF}.\\

 In Homer's Iliad, many people accused Zeus of \say{wanting to give victory to the Trojans.} But Zeus chose to remain \textbf{ALOOF}: \say{He sat apart in his all-glorious majesty, looking down upon the Trojans, the ships of the Achaeans, the gleam of bronze, and alike upon the slayers and the slain.}
 
\item \textbf{TRITE, HACKNEYED, BANAL, PLATITUDINOUS, INSIPID}\\ 
\textit{Unoriginal; commonplace; overused; \textbf{CLICHED}}\\

$\bullet$ In The Catcher in the Rye, Holden Caulfield just can't help seeing most people as \say{phony}-his favourite word. When he goes to hear Ernie, the jazz piano player, he thinks of the playing as \textbf{BANAL}: so lacking in originality that it is almost boring. He sees straight through his headmaster’s \textbf{PLATITUDE} that \say{Life is a game,} understanding the message to be \textbf{TRITE}, unoriginal, and lacking freshness. Many people who read The Catcher in the Rye today think of Holden Caulfield’s very character as \textbf{HACKNEYED}, because he represents a character we have seen all too many times: the moody, \textbf{DISAFFECTED} (disconnected), disgruntled teenager. But back in 1951, when the novel was first published, Salinger's portrait of a young person was considered \textbf{SEARINGLY} (scorchingly) original. Paula Abdul, the former American Idol and X-Factor judge, was known for being nice and \textbf{AFFABLE} (Word 18), always saying something positive to the contestants. Although Paula was nice, her comments were \textbf{TRITE}, \textbf{BANAL}, and \textbf{HACKNEYED}. According to \textbf{PLATITUDINOUS} Paula, every singer was \say{great,} \say{beautiful,} and \say{amazing.} She encouraged each one with pleasant but \textbf{INSIPID} compliments like \say{You're authentic,} \say{America loves you,} and \say{Your journey of magic is just beginning.}

\item \textbf{ANTECEDENT}\\
\textit{A preceding event; a \textbf{FORERUNNER}; a \textbf{PRECURSOR} gnify.}\\  

$\bullet$ Many critics have noted that the 1995 Disney movie Pocahontas can be viewed as a thematic \textbf{ANTECEDENT} to the 2010 blockbuster Avatar.  In Pocahontas, \textbf{AVARICIOUS} (Word 255) English settlers search for gold. In Avatar, an \textbf{AVARICIOUS} company wants to mine unobtanium from the fictional planet Pandora. In both movies beautiful \textbf{INDIGENOUS} (Word 47) women rescue soldiers who find themselves drawn to the native peoples they originally intended to conquer. By helping Captain John Smith discover the New World's life and beauty, Pocahontas serves as an \textbf{ANTECEDENT} for Avatar’s Neytiri.

\begin{tcolorbox}
\begin{center}
\textbf{ KNOW YOUR ROOTS}\\
\line(1,0){250}\\
\textbf{GREEK ROOT: ANTE }$ -$ \textit{before.}\\ 
\end{center}  
 
\textbf{ANTEBELLUM} $ -$ \textit{before the Civil War}\\ \textbf{ANTEDILUVIAN} $ -$ \textit{before the Biblical flood, a hyperbolic word describing something extremely old.}\\ \textbf{ANTEDATE} $ -$ \textit{to precede in time.}\\
\textbf{ANTEROOM} $ -$ \textit{a waiting room outside a larger room.}\\ 
\textbf{ANTERIOR} $ -$ \textit{before in time and place.}
\end{tcolorbox}

\item \textbf{PLAUSIBLE}\\ \textit{Believable; credible}\\ 
\textbf{IMPLAUSIBLE}\\
\textit{Unbelievable; incredible}\\ 

$\bullet$ Let's play \textbf{PLAUSIBLE} or \textbf{IMPLAUSIBLE}: In the Bourne Ultimatum, Jason Bourne successfully breaks into Noah Vosen's heavily-guarded top- security office and steals an entire set of classified Blackbriar documents. \textbf{PLAUSIBLE} or \textbf{IMPLAUSIBLE}? \textbf{PLAUSIBLE}—because he is Jason Bourne! I n The Avengers, Iron Man, Captain America, Thor, The Hulk, Hawkeye, and Black Widow successfully save New York City from an extraterrestrial attack and a nuclear missile. \textbf{PLAUSIBLE} or \textbf{IMPLAUSIBLE}? \textbf{PLAUSIBLE}-because The Avengers all have special skills and powers that allow them to defeat their foes!

\item \textbf{PRUDENT}\\
\textit{Careful; cautious; sensible}\\

$\bullet$ In the Twilight \textbf{SAGA} (Word 236), Bella Swan is a high school student who meets and falls in love with Edward Cullen. However, Edward is not just another high school student. He is a 107-year-old vampire who stopped aging physically at 17. Edward understands that their relationship poses grave dangers to Bella. However, Bella and Edward love each other, so they decide to stay together despite the danger. Together, they must be \textbf{PRUDENT} in dealing with the dangers that they face, among them werewolves, vengeful vampires, and the \textbf{OMINOUS} (Word 197) Volturi

\item \textbf{AESTHETIC}\\
\textit{Relating to the nature of beauty, art, and taste; having a sense of what is beautiful, attractive, or pleasing.}\\

$\bullet$ Do you know why the Mona Lisa is considered one of the most beautiful paintings of all time? The answer lies in its use of the Golden Ratio, the naturally occurring ratio of height to width that is most \textbf{AESTHETICALLY} pleasing to humans. The Mona Lisa's face is composed entirely of Golden Ratio rectangles and thus adds to the overall \textbf{AESTHETIC} of the painting. However, the Golden Ratio is not limited to art. Examples can be found in ancient Greek architecture, Egyptian pyramids, biology, and even widescreen television screens! It is not \textbf{AESTHETICALLY} pleasing if a character introduced at the very end solves a novel or play's conflicts. Aristotle criticized Euripides' play Medea for having Medea saved at the end by a character not integral to the plot. To his mind, \textbf{AESTHETICALLY} this was not a satisfying conclusion.

\item \textbf{PARADOX}\\ 
\textit{A seemingly contradictory statement that nonetheless expresses a truth}\\

$\bullet$ One of the most famous literary first lines is that of Charles Dickens's A Tale of Two Cities : \say{It was the best of times, it was the worst of times.} How could such a contradiction be true? In the course of the book, this \textbf{PARADOXICAL} statement is shown to be valid. In Mary Shelley's novel Frankenstein, the creature encounters many \textbf{PARADOXES}. One is the simultaneous positive and negative characteristics of fire. It can warm him, protect him, light his way, and cook his food, but it can also burn and destroy. Similarly, the creature also comes to recognize the \textbf{PARADOXICAL} nature of man: driven by conflicting forces of selfishness and \textbf{ALTRUISM} (Word 20).
\item \textbf{ENIGMATIC, INSCRUTABLE}\\ \textit{Mysterious; puzzling; unfathomable; baffling}\\ 

$\bullet$What do Da Vinci's Mona Lisa, Fitzgerald's description of Jay Gatsby, and J.K. Rowling's portrayal of Snape have in common? All three figures are \textbf{ENIGMATIC}. The Mona Lisa's \textbf{ENIGMATIC} smile has puzzled art lovers for centuries. When The Great Gatsby opens, Jay Gatsby is an \textbf{ENIGMATIC} figure whose great wealth and extravagant parties spark endless gossip. And Snape's personality and loyalties remain \textbf{INSCRUTABLE} until the final chapters of Harry Potter and the Deathly Hallows. 
\item \textbf{ACQUIESCE}\\ \textit{To comply; agree; give in.}\\ 

$\bullet$ In Pirates of the Caribbean: The Curse of the Black Pearl, Elizabeth Swann and Captain Barbossa conduct negotiations that include long words. Elizabeth Swann: Captain Barbossa, I am here to negotiate the cessation of hostilities against Port Royal. Captain Barbossa: There be a lot of long words in there, Miss. We're naught but humble pirates. What is it that you want? Elizabeth Swann: I want you to leave and never come back. Captain Barbossa: I'm disinclined to \textbf{ACQUIESCE} to your request. Means no!  Although he is a \say{humble pirate,} Captain Barbossa can use long words as well as she can.

\item \textbf{NAÏVE, GULLIBLE}\\ 
\textit{simple; lacking worldly expertise; overly \textbf{CREDULOUS}; unsophisticated; immature; inexperienced; \textbf{INGENUOUS} (Word 428)}\\  

$\bullet$ Nemo, of Finding Nemo, is a young clown fish who thinks he is old enough to swim out in the open waters. Young, \textbf{NAÏVE}, and wanting to defy his overprotective father, he wanders too near a boat. Suddenly, a net surrounds him. He is taken aboard the boat and from there to Sydney, Australia, to live in a fish tank. His father Marlin, \textbf{DESPONDENT} (Word 176) at his loss, vows to find his son. Marlin succeeds and ultimately brings Nemo back home. By the end of the film, Nemo has learned the importance of obeying his father and of not being so \textbf{GULLIBLE}.

\item \textbf{AUTONOMY}\\ \textit{Independence; self-governance} \\ \textbf{AUTONOMOUS}\\ \textit{Acting independently, or having the freedom to do so; not controlled by others.}\\ 

$\bullet$ Fahrenheit 451, the classic novel by Ray Bradbury, imagines a \textbf{DYSTOPIA} (an imaginery society characterized by oppression and human misery) society in which a faceless government exerts huge control over its citizens. No books are allowed; instead, citizens watch endless television streams of \textbf{PROPAGANDA} (zealous advancement of a group's principles) from the government. Bradbury's novel suggests that people naturally den \textbf{AUTONOMY} in their own lives; if a faceless government tries to exert authority over them, they will tend to be \textbf{SUBVERSIVE} (tending to overthrow), and rebel against that authority. In the movie Men in Black, Agent Zed explains that MIB is an \textbf{AUTONOMOUS} organization that is \say{not a part of the system.} He goes on to say that MIB is \say{above the system, over it, beyond it; we are they, we are them, we are the Men in Black.} They are serious about their \textbf{AUTONOMY}.
\item \textbf{FUTILE}\\ \textit{Completely useless; doomed to failure; in vain.}\\

$\bullet$ The Deepwater Horizon oil spill released a \textbf{PRODIGIOUS} (huge, massive) flood of crude oil into, the Gulf of Mexico. BP engineers made repeated attempts to control or stop the spill. However, all of their initial efforts proved to be \textbf{FUTILE}. Although crews worked tirelessly to protect hundreds of miles of beaches, wetlands, and estuaries, local residents worried that these efforts would also prove to be \textbf{FUTILE}.
\item \textbf{INDIGENOUS, ENDEMIC}\\ \textit{Native to an area}\\ 

$\bullet$ Which of the following are Old World plants and animals, and which are New World plants and animals: potatoes, tomatoes, maize, sunflowers, cocoa beans, turkeys, and buffaloes? Surprisingly, all of these plants and animals are \textbf{INDIGENOUS} or \textbf{ENDEMIC} to the New World.
\begin{tcolorbox}
\begin{center}
\textbf{ KNOW YOUR ROOTS}\\
\line(1,0){250}\\
\textbf{GREEK ROOT: DEM, DEMO} $ - $ \textit{ the people.}
\end{center}  
  \textbf{PANDEMIC} (Word 49) $ - $  \textit{of all the people, prevalent over a whole area.}\\ 
\textbf{DEMOCRACY} $ - $ \textit{rule by the people, by the majority.}\\ 
\textbf{DEMAGOGUE} (Word 111)$ - $ \textit{a person who tries to stir up the people by appealing to emotion and prejudice in order to achieve selfish ends.}\\
 \textbf{DEMOGRAPHICS} (Word 433) $ - $ \textit{the science of vital statistics about populations (births, deaths, marriages, incomes, etc.)}\\ 
\textbf{EPIDEMIC} $ - $ \textit{a rapid spread of a contagious disease or other negative condition.}
\end{tcolorbox}
\item \textbf{UBIQUITOUS, PREVALENT}\\ \textit{Characterized by being everywhere; omnipresent; widespread}; \textbf{PERVASIVE}.\\ 

$\bullet$ What do cell phones, iPods, Starbucks The coffee shops, and McDonald's fast-food restaurants have in common? They are all \textbf{UBIQUITOUS}-we see them everywhere. Popular fashions are also \textbf{PERVASIVE}. For example, baggy knee-length shorts have completely replaced the once-\textbf{PREVALENT} short shorts of the 1970s. From high school b-ballers to WNBA and NBA superstars, long shorts are now \textbf{UBIQUITOUS}.
\item \textbf{PANDEMIC} -  \textit{An epidemic that is geographically widespread and affects a large proportion of the population.}\\ 

 $\bullet$ In the movie I Am Legend, a manmade virus known as KV triggers a global \textbf{PANDEMIC} that kills almost all of the human population on Earth. While there has never been a real \textbf{PANDEMIC} of this magnitude, virus strains and diseases have caused widespread deaths. In 1347 the Black Plague killed as many as one-third of the people in Europe. In the 16th century, Spanish conquistadores spread small pox and other diseases that \textbf{DECIMATED} (destroyed a great proportion of) the \textbf{INDIGENOUS} (Word 47) populations in Central America, the Caribbean, and Mexico. Our own times have not been immune to epidemics. The 1918 flu \textbf{PANDEMIC} killed 50 to 100 million people, and more recently we have had SARS, Asian Bird Flu, and Swine Flu \textbf{PANDEMICS}.\\ \textbf{PANDEMIC} can also be used as an adjective, meaning \textbf{PREVALENT} (Word 48) over a large area.
\item \textbf{FORTITUDE}\\  \textit{Strength of mind that allows one to endure pain or adversity with courage}.\\ 

$\bullet$ William Lloyd Garrison and Rosa Parks demonstrated great personal \textbf{FORTITUDE}. While most Americans accepted slavery, Garrison boldly demanded the immediate and unconditional emancipation of all slaves. Although initially ignored, Garrison \textbf{PERSEVERED} (refused to give up no matter the situation) and lived to see President Lincoln issue the Emancipation Proclamation. Rosa Parks also illustrates the principle that \textbf{FORTITUDE} is needed to achieve difficult goals. While most Americans accepted segregation, Rosa refused a bus driver's order to give up her seat to a white passenger. Her historic action helped \textbf{GALVANIZE} (Word 148) the Civil Rights Movement.
\end{enumerate}
\chapter{Core Vocabulary (51-100)}
Chapter 2 continues to build the list of 100 Core Vocabulary Words. As in Chapter 1, each of these words has been the key to a Level 3 or Level 4 question. We \textbf{EXHORT} (Word 53) you to study hard. As always, our \textbf{PENCHANT} (Word 62) for vivid pop culture examples will help you learn and remember new words. So don't let the Core Words \textbf{THWART} (Word 67) you. Now is the time to \textbf{TENACIOUSLY} (Word 56) pursue your goal of conquering the SAT. Remember, there is \textbf{INCONTROVERTIBLE} (Word 70) proof that your Critic heal Reading score will go up as your vocabulary goes up!
\begin{enumerate}
 \setcounter{enumi}{50}
\item \textbf{DIMINUTIVE}\\
\textit{Very small}\\

$\bullet$ The 2012 Summer Olympics coverage featured the latest team of \textbf{DIMINUTIVE} gymnasts, five American girls who averaged 16 years old, 5'1'' tall, and 104 pounds, and who performed astounding feats of strength, agility, and precision. Led by Gabby Douglas, who also took first place in the individual all-around event, the \textbf{DIMINUTIVE} Fab Five brought home the women’s team gold medal for the first time since 1996. Gabby, at 4’11’ and 90 pounds the most \textbf{DIMINUTIVE} of the bunch, was dubbed  \say{the flying squirrel} by Marta Karolyi, the U.S. national team coordinator. Given the rigors of the intense training, the high incidence of injuries,and the other sacrifices required in order to reach the highest level of the sport, women gymnasts have a short shelf life. But in August 2012, \textbf{DIMINUTIVE} Gabby Douglas won the hearts of Americans everywhere.
\item \textbf{TRIVIAL}\\
 \textit{ unimportant; insignificant}\\
 \textbf{MINUTIAE}\\
 \textit{Minor everyday details}\\
 
 $\bullet$ Drake is one of the world's most popular hip hop artists. While Drake would prefer to concentrate on creating music, his zealous fans often focus on interesting but \textbf{TRIVIAL MINUTIAE} about his personal life. For example, Drake was raised by a Jewish mother and had a Bar Mitzvah. And online rumors continue to link him with Rihanna!

\item \textbf{EXHORT}\\
\textit{ To encourage; urge; give a pep talk;}\textbf{ IMPLORE}\\

$\bullet$ American League baseball player Derek Jeter has spent his entire career with the New York Yankees. Naturally, New York fans love him. When Jeter began to approach the \textbf{COVETED} (Word 32) 3,000 hit milestone, his teammates and fans \textbf{EXHORTED} him to continue to play well so he could reach the \textbf{ELUSIVE} (Word 161) milestone. The \textbf{EXHORTATIONS} worked. On July 9, 2011 he became the first New York Yankee to reach the 3,000 hit mark. Even sweeter, his 3,000th hit was a home run! 

\item \textbf{ANTIPATHY}\\
\textit{Strong dislike; ill will; the state of \textbf{DETESTING} someone; \textbf{ENMITY; RANCOR} }\\

$\bullet$ In The Social Network, League Cameron and Tyler Winklevoss and their business partner, Divya Narendra, approached Mark Zuckerberg with an idea they called \say{HarvardConnection,} an online social network exclusively for Harvard University students and alumni. Zuckerberg broke his agreement with the HarvardConnection team and approached his friend Eduardo Saverin about a nearly identical website idea called \say{The facebook.} The first website to allow the entire campus to communicate and socialize with ease, Facebook skyrocketed in popularity.\\

Can you imagine the \textbf{ANTIPATHY} that the Winklevoss brothers and Narendra felt upon seeing their idea becoming successful without them? In the movie scene, Narendra discovered Zuckerberg's Thefacebook, slammed his laptop closed, and, filled with \textbf{ENMITY}, stormed out of the room to inform the Winklevoss brothers of Zuckerberg's betrayal.

\item \textbf{DIGRESS}\\
\textit{To depart from a subject; wander; ramble.}\\

$\bullet$ Have you ever listened to someone who repeatedly wanders off a topic? If so, then you know how confusing and annoying it is when a speaker \textbf{DIGRESSES} from a subject. In the movie Office Space, Milton is \textbf{NOTORIOUS} (widely but unfavorably known) for his long-winded \textbf{DIGRESSIONS}. \textbf{DIGRESSING} is not limited to speaking. Writers sometimes \textbf{DIGRESS} or wander off a topic. On the SAT I, your first task will be to write an essay. Readers reward essays that are well-organized and deduct points from essays that \textbf{DIGRESS} from the topic.

\begin{tcolorbox}
\begin{center}
\textbf{ KNOW YOUR ROOT}\\
\line(1,0){250}\\
\textbf{LATIN ROOTS: GRESS} $ - $ \textit{to step}.\\  
\end{center}  
\textbf{PROGRESS} $ - $ \textit{to step forward.}\\
\textbf{REGRESS} $ - $ \textit{to step back.}\\
\textbf{TRANSGRESS} $ - $ \textit{ to step across the line that divides right from wrong.}\\
\textbf{EGRESS} $ - $ \textit{to step out, to exit (or as a noun, an exit).}\\
\textbf{AGGRESSIVE} $ - $ \textit{tending to attack, encroach, or step on others.}
\end{tcolorbox}

\item \textbf{TENACIOUS}\\
\textit{Characterized by holding fast to something valued; showing great determination.}\\

$\bullet$ The 2006 film Amazing Grace tells the story of the 20-year campaign against the British slave trade led by the \textbf{TENACIOUS} Member of Parliament William Wilberforce. Through Wilberforce's \textbf{TENACITY} and determination, the battle \textbf{CULMINATED} (reached completion) in the 1807 bill that abolished the slave trade in the British Empire. Though the film is not totally accurate historically, itidt TCU does make the inspiring point of the effectiveness of sheer \textbf{TENACITY} in the face of almost impossible odds.\\

Think too about the \textbf{TENACITY} of a dog with a bone, never letting go. That might remind you of the word \textbf{DOGGED} (pronounced with two syllables), which means having the \textbf{TENACITY} of a dog. You can also use \textbf{TENACIOUSNESS} instead of \textbf{TENACITY}; they mean the same thing.
\begin{tcolorbox}
\begin{center}
\textbf{Tip for a Direct Hit}\\
\line(1,0){250}\\
\end{center}  
The root of \textbf{TENACIOUS} is the Latin root \textbf{TEN}, \say{to hold.} You can find it in \textbf{TENET}, an opinion, idea, or principle \textbf{HELD} true by a person or organization. It's also in \textbf{TENABLE}, which means capable of being \textbf{HELD}, defended, and logically supported, as in a \textbf{TENABLE} argument or thesis.
\end{tcolorbox}

\item \textbf{INDULGENT}\\
\textit{Characterized by excessive generosity; overly tolerant}\\

$\bullet$ In the movie Mean Girls, Regina George's mother prides herself on being \textbf{INDULGENT}. She proudly tells Regina and Cady, \say{I just want you to know, if you ever need anything, don't be shy, OK? There are NO rules in the house. I'm not like a \say{regular} mom. I'm a \say{cool} mom.} Mrs. George should have said, \say{I'm a super-\textbf{INDULGENT} mom who lets Regina do anything she wishes}.

\item \textbf{POLARIZE}\\
\textit{To create disunity or dissension; to break up into opposing factions or groups;\\ to be \textbf{DIVISIVE}.}\\

$\bullet$ Americans have a long and distinguished record of settling differences by reaching a compromise. However, some issues are so \textbf{DIVISIVE} and \textbf{POLARIZING} that a compromise is impossible.  Before the Civil War, the issue of slavery \textbf{POLARIZED} Americans into two groups: those who defended the South's \say{peculiar institution} and those who demanded that slavery be abolished. As Lincoln eloquently noted: \say{A house divided against itself cannot stand. I believe this government cannot endure permanently half slave and half free.}

\item \textbf{NEBULOUS}\\ 
\textit{Vague; cloudy; misty; lacking a fully developed form.}\\

$\bullet$ Have you read the Epilogue in Harry Potter and the Deathly Hallows? If you found it rather vague, then J.K. Rowling achieved her goal. In an interview, Rowling stated that the Epilogue is deliberately \say{\textbf{NEBULOUS.}} She wanted readers to feel as if they were looking at Platform $9\dfrac{3}{4}$ through the mist, unable to make out exactly who was there and who was not.

\item \textbf{ANALOGY}\\
\textit{ A similarity or likeness between things-events, ideas, actions, trends-that are otherwise unrelated.}\\
\textbf{ANALOGOUS}\\Comparable or similar in certain respects.\\
 
$\bullet$ Did you know that for most of its history the SAT included a number of \textbf{ANALOGY} questions? For example, students were asked to see the \textbf{ANALOGY} or similarity between a tree and a forest and a star and a galaxy. The \textbf{ANALOGY} is that a tree is part of a forest in the same way that a star is part of a galaxy. Although the College Board removed analogies in 2005, SAT test writers still expect students to recognize \textbf{ANALOGIES} in critical readings. Don't be confused by the phrase \say{is most \textbf{ANALOGOUS} to.} The question is asking you to identify a situation or example that is most similar to the one in the reading passage.

\item \textbf{EPHEMERAL, FLEETING, EVANESCENT}\\ 
\textit{Very brief; lasting for a short time; transient.}\\ 

\textbf{PERENNIAL}\\
\textit{ Returning year after year; enduring}\\

$\bullet$ What do the following groups and their hit songs have in common: \say{Who Let the Dogs Out?} by Baha Men, \say{Stuck In The Middle With You} by Stealers Wheel, and \say{It's Raining Men} by the Weather Girls? All three groups were \say{one-hit wonders} who had a single hit song and then disappeared. Their popularity was \textbf{EVANESCENT}. They were \textbf{EPHEMERAL}-here today and gone tomorrow. On the contrary, bands like The Beatles, The Beach Boys, and Simon \& Garfunkel have remained \textbf{PERENNIAL} favorites. The Beatles' albums continue to be bestsellers on iTunes. The Beach Boys still maintain a busy tour schedule, and the songs of Simon \& Garfunkel remain staples of popular culture. Paul Simon was even asked to perform their hit song \say{The Sound of Silence} at the 9/11 tenth anniversary memorial service. All three of these bands have maintained immense popularity throughout the decades.

\item \textbf{PENCHANT, PREDILECTION, PROPENSITY}\\
\textit{ A liking or preference for something; a \textbf{BENT} (Word 299); an \textbf{INCLINATION}}\\

$\bullet$ What do film star Angelina Jolie and rap artist Lil Wayne have in common? Both have a well-known \textbf{PENCHANT} for tattoos. Angelina's tattoos include a prayer of Buddhist Sanskrit symbols to honor her first adopted son, Maddox, coordinates representing the geographic neddddntain ave locations of her children's birthplaces, and the statement \say{know your rights.} Lil Wayne's \textbf{PREDILECTION} for tattoos has led him to cover his face and torso with tattoos. For example, a red tattoo above his right eyebrow states, \say{I am music,} emphasizing his love of music. The numbers 9 27 82 on his right forearm are his date of birth.
 \item \textbf{CAPRICIOUS, MERCURIAL, FICKLE}\\
 \textit{ Very changeable; characterized by constantly-shifting moods}\\
 
 $\bullet$ When the gossip magazine Us Weekly  published a story and photos of Twilight's Kristen Stewart cheating on her boyfriend Robert Pattinson, passionate Twilight fans responded in disbelief on their Twitter accounts. The outraged fans \textbf{LAMBASTED}. (Word 310) the magazine and insisted that the photos of Stewart were fake. However, Stewart released a statement apologizing for her \say{momentary indiscretion} and declaring her love for Pattinson. The \textbf{FICKLE} fans turned \textbf{CAPRICIOUSLY} from supporting Stewart to \textbf{CASTIGATING} (severely criticizing ) her and mourning the end of their favorite celebrity couple. Fans added \textbf{MAUDLIN} (Word 142) videos and social media posts online, \textbf{BEMOANING} (to express grief over) the breakup of \say{Robsten.}  
\begin{tcolorbox}
\begin{center}
\textbf{ Medieval Humours}\\
\line(1,0){250}\\
\end{center}  
  In medieval times, it was believed that people's personalities or moods were determined by the relative amounts of the four bodily fluids (or \textbf{HUMOURS}) in their bodies. Though we no longer believe in the physiological basis, we still use the words to describe people. 
 \begin{center}
\centering%
\begin{tabular}{p{\dimexpr0.30\textwidth-2\tabcolsep-\arrayrulewidth\relax}
                p{\dimexpr0.25\textwidth-2\tabcolsep-\arrayrulewidth\relax}
                p{\dimexpr0.45\textwidth-2\tabcolsep-\arrayrulewidth\relax}
              }
\hline
\textbf{Predominant Fluid} & \textbf{Temperament} &  \textbf{Aspects } \\

Blood & \textbf{SANGUINE} & cheerful, hopeful, optimistic\\
\hline
Black bile & \textbf{MELANCHOLY} & gloomy, depressed, \textbf{DESPONDENT}, \textbf{PENSIVE}  \\
\hline
Yellow bile & \textbf{CHOLERIC} &  angry, irritable, \textbf{IRASCIBLE}\\
\hline
Phlegm & \textbf{PHLEGMATIC} & self-possessed, imperturbable, calm, \textbf{APATHETIC}, sluggish \\
\hline
Fluctuating among all four fluids & \textbf{MERCURIAL} & volatile, changeable, \textbf{FICKLE}.\\
\hline 
\end{tabular}
\end{center}
\end{tcolorbox}
                                                                                  
\item \textbf{BOORISH, UNCOUTH, CRASS}\\
\textit{Vulgar; characterized by crude behavior and deplorable manners; unrefined}\\

$\bullet$ Billy Madison ( Billy Madison), Ron Burgundy ( Anchorman), Borat ( Borat), and Ben Stone ( Knocked Up) all demonstrated \textbf{BOORISH} manners and behaviors. However, none of these \textbf{UNCOUTH} characters quite equaled Bluto in Animal House. In a classic scene, Bluto piled food onto h" tttt Tis cafeteria plate while stuffing food in his pockets. He then sat down uninvited at a cafeteria table. Disgusted by Bluto's outrageous appearance and \textbf{CRASS} manners, Mandy called him a \say{P-I-G, pig.} Undeterred by Mandy's insult, Bluto stuffed mashed potatoes into his mouth and asked Mandy and her \textbf{INCREDULOUS} (Word 363) friends, \say{See if you can guess what I am now.} He then pressed his hands against his cheeks, causing the mashed potatoes to spray onto the shocked diners. Pleased with his \textbf{BOORISH} antics, Bluto proudly answered his own question by announcing, \say{I'm a zit! Get it?}
 \item \textbf{INDIGNANT}\\
\textit{ Characterized by outrage at something that is perceived as unjust}\\

$\bullet$ What do Andrew Jackson's supporters in 1824 and Al Gore's supporters in 2000 have in common? Both were \textbf{INDIGNANT} at the outcomes of presidential elections. Following the election of 1824 Andrew Jackson's \textbf{INDIGNANT} supporters accused John Quincy Adams and Henry Clay of stealing the election from Old Hickory. Following the election of 2000, Al Gore's \textbf{INDIGNANT} supporters accused George W. Bush and the U.S. Supreme Court of stealing the election from Gore. 
\item \textbf{INNUENDO}\\
\textit{A veiled reference; an insinuation}\\

$\bullet$ At the beginning of The Godfather,  Kay does not understand the workings of the Corleone family business, and she asks Michael how his father managed a business deal. Michael responds with an \textbf{INNUENDO}: \say{My father made him an offer he couldn't refuse.} His response insinuates that Don Vito uses coercion and threats in his business dealings. Michael's \textbf{INNUENDO} suggests that his father is a powerful mob boss. 

\item\textbf{ THWART, STYMIE}\\
 \textit{To stop; to frustrate; to prevent}\\
 
 $\bullet$ In the Harry Potter SAGA (Word 236) Lilly Potters love \textbf{THWARTED} Lord Voldemort's attempt to kill her one-year-old son, Harry. With the help of Ron and Hermione, Harry repeatedly THWARTED the Dark Lord's attempts to kill him.  At the Paris Peace Conference at the end of World War I, which \textbf{CULMINATED} (concluded) in the Treaty of Versailles, most of President Woodrow Wilson's proposals for a \say{Just Peace} were \textbf{THWARTED} by the other world leaders, who were more interested in \textbf{RETRIBUTION} (punishment, vengeance). They did approve his plan for a League of Nations, which he hoped would be able to prevent future wars. When Wilson presented the treaty to the U.S. Senate, there was much opposition. The treaty went down to defeat, Wilson's efforts were again \textbf{STYMIED}, and the weak League of Nations never achieved its goals, lacking the participatihe tton of the world’s newest superpower.  After the \textbf{CHICANERY} (Word 339) that came to light in 2011, new security measures have been implemented by both the SAT and the ACT to \textbf{STYMIE} those who might wish to take an exam for someone else. Students will now be required to upload a photo of themselves when registering for these exams. The photos will appear on the students' admission tickets and on the test site rosters available to proctors. Then the photos will be attached to any score reports sent to high schools and colleges. 
 
\item \textbf{ADROIT, DEFT, ADEPT}\\
\textit{ Having or showing great skill; \textbf{DEXTEROUS}; nimble}\\

$\bullet$ What do 16-year-old Austin Wierschke and action star Chuck Norris have in common? Austin has \textbf{DEXTEROUS} hands, and Chuck has \textbf{ADROIT} legs. Austin won the U.S. National Texting Championship two years in a row. He beat out 11 other finalists by \textbf{DEFTLY} texting blindfolded, texting with his hands behind his back, and by enduring rounds of marathon texting. As everyone knows, Chuck Norris is \textbf{ADEPT} at using a roundhouse kick to escape even the toughest situations. In fact, it is rumored that if someone were DEFT enough to harness the energy from a Chuck Norris roundhouse kick, he or she could power the entire country of Australia for 44 minutes. 

\begin{tcolorbox}
\begin{center}
 \textbf{Tip for a Direct Hit }\\
 \line(1,0){250}\\
Are you right-handed or left handed? Right-handed people were once thought to be more \textbf{ADROIT} an     \textbf{DEXTEROUS} than left-handed people. This bias can be seen in the etymology of these two words. The English word \textbf{ADROIT} is actually derived from the French word \textit{droit} meaning right, as opposed to left. So if you are \textbf{MALADROIT}, you are not skillful. The ancient Romans shared the same positive view of right-handed people. The Latin word \textit{dexter} means right, as opposed to left.   

\end{center}
\end{tcolorbox} 


\item \textbf{ADMONISH}\\
\textit{ To earnestly caution; to warn another to avoid a course of action}\\

$\bullet$ First sung in November 1934, \say{Santa Claus is Coming to Town} celebrates Santa's much-anticipated arrival on Christmas Eve. However, while Santa may be very \textbf{MUNIFICENT} (Word 258), he is also very \textbf{VIGILANT} (watchful, alert). He keeps a list, and he knows \say{who's naughty or nice.} The song earnestly \textbf{ADMONISHES} children to \say{be good for goodness sake.}
\begin{tcolorbox}
\begin{center}
 \textbf{KNOW YOUR ROOTS}\\
 \line(1,0){250}\\
  \textbf{ LATIN ROOT: MON} $ - $ \textit{ to warn, remind} 
\end{center}
 \textbf{ADMONITION} $ - $ \textit{ a warning or reproof, a reminder}\\ \textbf{PREMONITION} $ - $ \textit{ a warning in advance, \textbf{PRESENTIMENT} (hint) of something evil, foreboding}\\
\textbf{MONITOR} $ - $  \textit{ a person or a device that reminds or checks (like a study hall monitor, a heart  monitor, or an audio monitor for performers on a stage)} \\
\textbf{MONUMENT} $ - $ \textit{ a sepulchre, memorial , edifice to commemorate something or someone notable,  something that reminds (literally)}
\end{tcolorbox} 
 
\item \textbf{INCONTROVERTIBLE}\\
\textit{Impossible to deny or disprove; demonstrably true}\\

$\bullet$ In recent years the global warming debate has grown increasingly heated (no pun intended), politicized, and \textbf{POLARIZED} (Word 58). Al Gore's film, An Inconvenient Truth, presented statistics that many people challenge. But it is becoming clear that global warming is an \textbf{INCONTROVERTIBLE} fact. What is less clear has been the cause of the climatic changes. Many \textbf{CONCEDE} (admit) the existence of the trend but claim that the current trend is merely part of a natural \textbf{METEOROLOGICAL} (having to do with weather) cycle. Others lay the blame on humans' emission of greenhouse gases. According to Richard A. Muller, a former \textbf{SKEPTIC} (Word 102) whose Berkeley Earth Surface Temperature project has persuaded him of human \textbf{CULPABILITY} (blameworthiness) in global warming, the changes are too great to be \textbf{ASCRIBED} (attributed) to urban heating, solar activity, world population, normal fluctuations, or manipulation of data. Only changes in the carbon dioxide curve match the changes in world temperatures. So, the \textbf{SKEPTIC} has been persuaded that man is, \textbf{INCONTROVERTIBLY}, playing a part in the climate changes we are now experiencing.

\item \textbf{VORACIOUS, RAVENOUS, RAPACIOUS}\\
\textit{  Having a huge appetite that cannot be satisfied; \textbf{INSATIABLE}}\\

$\bullet$ What do Homer (The Simpsons), Bluto (Animal House), and Scooby-Doo ( Scooby Do! Mystery Incorporated) have in common? All three have \textbf{VORACIOUS} appetites. Homer has an \textbf{INSATIABLE} appetite for frosted doughnuts. Bluto regularly and \textbf{RAPACIOUSLY} piles great quantities of food on his plate. Scooby has a \textbf{RAVENOUS} appetite for Scooby Snacks, and he habitually sneaks food from the plates of his friends. 

\item \textbf{CALLOUS}\\
\textit{ Emotionally hardened; insensitive; unfeeling.}\\

$\bullet$ In the movie Mean Girls, the Plastics \textbf{CALLOUSLY} mistreat their classmat appppp D:es. They even keep a \say{Burn Book} filled with \textbf{CALLOUS} \textbf{INNUENDOES} (Word 66) and \textbf{SARCASTIC} (Word 3) putdowns.  In F. Scott Fitzgerald's novel The Great Gatsby, Tom Buchanan \textbf{CALLOUSLY} ruins the lives of four people (Daisy, Gatsby, Myrtle, and George) while recklessly pursuing his own selfish pleasures.

\item \textbf{INTREPID, UNDAUNTED}\\
\textit{ Courageous, \textbf{RESOLUTE} (Word 359), and fearless}\\

$\bullet$ What do Luke Skywalker and Charles Lindbergh have in common? Both were \textbf{INTREPID} pilots who were \textbf{UNDAUNTED} by seemingly impossible odds. In the movie Star Wars: Episode IV, Luke was \textbf{UNDAUNTED} by the Empire's seemingly invincible Death Star. The \textbf{INTREPID} Skywalker destroyed the Death Star with well-aimed proton torpedoes. The American aviator Charles Lindbergh was also \textbf{UNDAUNTED} by a seemingly impossible task. Despite several attempts, no pilot had successfully flown across the Atlantic. In 1927, the \textbf{INTREPID} Lindbergh electrified the world by flying his single-engine plane, the Spirit of St. Louis, from New York to Paris in a grueling 33-hour and 39-minute flight. 

\item \textbf{NONCHALANT}\\
\textit{ Having an air of casual indifference; coolly unconcerned; \textbf{UNFLAPPABLE}}\\

$\bullet$ When you are driving, do you slow down for a yellow light and promptly stop for a red light? We hope so. While careful and law-abiding drivers follow these rules of the road, not all drivers do. Italian drivers are famous for their \textbf{NONCHALANT} attitude toward yellow and even red lights. One typical Italian cab driver \textbf{NONCHALANTLY} explained that lights are merely advisory: \say{Everyone drives through yellow lights and fresh red ones. It is no big deal.} Needless to say, we hope you will  not take such a \textbf{NONCHALANT} attitude.

\item \textbf{CONVOLUTED}\\
\textit{ Winding, twisting, and therefore difficult to understand; intricate}\\

$\bullet$  What do the Electoral College and the Bowl Championship Series (BCS) have in common? Both require a \textbf{CONVOLUTED} process to choose a winner. The Electoral College requires a \textbf{CONVOLUTED} process to choose a President, and the BCS requires a \textbf{CONVOLUTED} process to choose two football teams to play for the national championship. 

\item \textbf{ITINERANT}\\
\textit{Migrating from place to place; \textbf{NOT SEDENTARY}}\\

$\bullet$ During the Great Awakening, George Whitefield and other \textbf{ITINERANT} ministers touring the Colonies preached their message of human helplessness and divine \textbf{OMNIPOTENCE} (infinite power). Today, many movie stars also live \textbf{ITINERANT} lives. For example, during the last six years, Angelina Jolie and Brad Pitt have lived in 15 homes all over the world, including Paris, Prague, Los Angeles, New Orleans, Berlin, Namibia, India, and New York City. Jolie enjoys her \textbf{ITINERANT} lifestyle and says that it is important to experience a variety of cultures.

\item  \textbf{POIGNANT}\\
\textit{ Moving; touching; heartrending}\\

$\bullet$ In the movie Remember the Titans, Gerry Bertier and Julius Campbell are forced to become teammates on the racially-divided T.C. Williams High School football team. Although originally bitter rivals, they overcome their prejudices and become close friends. When Julius visits the paralyzed Gerry in the hospital, the nurse bars Julius, who is black, from the room, saying, \say{Only kin's allowed in here.} But Gerry corrects her: \say{Alice, are you blind? Don't you see the family resemblance? That's my brother.} This \textbf{POIGNANT} scene brought tears to the eyes of many viewers. 

\item \textbf{IMPETUS}\\
\textit{ A stimulus or encouragement that results in increased activity }\\

$\bullet$ Lord Voldemort's resurrection at the end of Harry Potter and the Goblet of Fire provided the \textbf{IMPETUS} for the revival of the Order of the Phoenix and the formation of Dumbledore's Army. Although it was a failure, Shays' Rebellion in 1786 alarmed key American colonial leaders, thus providing the \textbf{IMPETUS} for calling a convention to revise and strengthen the Articles of Confederation. 

\item \textbf{BUCOLIC, RUSTIC, PASTORAL}\\
\textit{ Characteristic of charming, unspoiled countryside and the simple, rural life}\\

$\bullet$  Americans have always been proud of our country's great natural beauty. During the early 19th century, a group of artists known as the Hudson River School specialized in painting the \textbf{RUSTIC} beauty of America's unspoiled lands cape. Today, many students are attracted to the \textbf{PASTORAL} beauty of campuses located in small towns. For example, one writer described Blacksburg, Virginia, the home of Virginia Tech, as \say{a quaint, off-the-beaten-track, \textbf{BUCOLIC} college town nestled in the mountains of southwest Virginia.}
 
 \item \textbf{EQUANIMITY}\\
 \textit{ Calmness; composure; even-temperedness; poise}\\
 
 $\bullet$ George Washington, the great Father of America, was known for his \textbf{EQUANIMITY}. He maintained composure no matter what happened around him. Faced with the dangers of battle during the Revolutionary War, Washington remained even-tempered and unflappable. His ability to maintain composure in the heat of battle encouraged his troops to follow and respect him, even during the most devastating times in the Revolution. His \textbf{EQUANIMITY} made him an indispensable leader in the early years of the fledgling nation.
 
\item \textbf{PANACHE, VERVE, FLAMBOYANCE, ÉLAN (Word 316)}\\ \textit{Great vigor and energy; dash, especially in artistic performance and composition }\\

$\bullet$ During the Middle Ages, proud European military commanders often placed feathers or a plume in their helmets as they rode into battle. Known as a panache, the feathers and plumes helped troops identify their commander but also made him an easier target for enemy arrows and bullets. Given the risk, it took real courage for a commander to wear a panache. Today the word \textbf{PANACHE} no longer refers to feathers or a plume. But \textbf{PANACHE} still retains its sense of \textbf{VERVE} or dash. \textbf{PANACHE} is now most frequently used to refer to \textbf{FLAMBOYANT} entertainers. For example, Lady Gaga is one of the music world's most \textbf{FLAMBOYANT} performers. 

\item \textbf{PROVOCATIVE}\\
\textit{ Provoking discussion; stimulating controversy; arousing a reaction}\\

$\bullet$ Prior to World War I, young women aspired to seem modest and maidenly. But that changed during the Roaring Twenties. Once \textbf{DEMURE} (modest) maidens now \textbf{PROVOCATIVELY} proclaimed their new freedom by becoming \say{flappers.} Flappers shocked their elders by dancing the Charleston and wearing one-piece bathing suits. Dismayed by this \textbf{PROVOCATIVE} clothing, officials at some beaches insisted on measuring the length of the bathing suits to make sure that they did not reveal too much of the women's legs. In today's world, this notion of \textbf{PROVOCATIVE} would seem \textbf{ARCHAIC} (Word 25)! 

\item \textbf{PLACID, SERENE}\\
\textit{ Calm or quiet; undisturbed by tumult or disorder}\\

$\bullet$ What do the Pacific Ocean and the SAT word \textbf{PLACID} have in common? When the legendary explorer Ferdinand Magellan left the Strait of Magellan, he entered an immense and as yet unexplored body of water that he describethaaaaaa tgd as a Mare Pacificum or \say{peaceful sea.} 
\begin{tcolorbox}
\begin{center}
 \textbf{KNOW YOUR ROOTS}\\
 \line(1,0){250}\\
  \textbf{ LATIN PREFIX: PLAC} $ - $ \textit{ to quiet, to soothe, pacify, please } 
\end{center}
 \textbf{IMPLACABLE} $ - $ \textit{ unappeasable, inexorable}\\
 \textbf{ PLACATE} $ - $ \textit{ to appease or calm someone's anger}\\     \textbf{PLACID} $ - $ \textit{ calm, quiet}\\
 \textbf{ COMPLACENT} $ - $ \textit{ self-satisfied, smug}\\
 \textbf{ COMPLAISANT} (Word 335) $ - $ \textit{ disposed to please (note French plaisir), affable, gracious}
\end{tcolorbox} 

 \item \textbf{FORTUITOUS}\\
 \textit{ Of accidental but fortunate occurrence; having unexpected good fortune}\\
 
$\bullet$ In the fall of 1862, the South appeared to be on the verge of victory in the Civil War. Following a brilliant triumph at the Second Battle of Bull Run, General Robert E. Lee boldly invaded Maryland. In war, however, decisive battles are often determined as much by \textbf{FORTUITOUS} accident as by carefully-planned strategy. As Lee's army st eadily advanced, a Union corporal discovered a bulky envelope lying in the grass near a shade tree. Curious, he picked it up and discovered three cigars wrapped in a piece of paper containing Lee's secret battle plans. This \textbf{FORTUITOUS} discovery played a key role in enabling the Union forces to win a pivotal victory at the Battle of Antietam. 

\item \textbf{DISPEL}\\
\textit{ To drive away; scatter, as to \textbf{DISPEL} a misconception}\\

$\bullet$ The first administration of the SAT occurred in 1901, and since then quite a few myths have arisen. We're here to \textbf{DISPEL} a couple of misconceptions you might have. Myth: The SAT is a reasoning test; you can't study for it. Absolutely not! The book you're holding right now, Direct Hits, can help you ace the sentence completions and boost your critical reading score. Myth: It's always better to leave a question blank than to guess on the SAT. Not necessarily. You receive a full point for correct answers, gain zero points for incorrect answers, and lose a quarter point for incorrect answers. But if you can eliminate one or two incorrect answers in a multiple choice, your odds of answering correctly improve dramatically. On the sentence completions section, if you see a Direct Hits word in the answer choices, but you're sure it's not the correct answer (an indirect hit), you can eliminate this answer and greatly improve your chance of guessing correctly To further \textbf{DISPEL} this misconception, remember what Wayne Gretzky says: \say{You miss 100 percent of the shots you don't take.} 

\item \textbf{AMALGAM}\\
\textit{ A mixture; a blend; a combination of different elements}\\

$\bullet$ Rap star Ludacris' name is actually an \textbf{AMALGAM}. He combined his birth name Cris and his radio handle Luda to \textbf{COIN} (Word 296) the new name-LUDACRIS! Similarly, rap star Jay-Z's name is also an \textbf{AMALGAM}. Shawn Carter grew up in Brooklyn near where the J-Z subway line has a stop on Marcy Avenue. Carter's friends nicknamed him \say{Jazzy.} Carter later combined the name of the subway line with his nickname to \textbf{COIN} the new name Jay-Z! 

\item \textbf{VIABLE, FEASIBLE}\\
\textit{ Capable of being accomplished; possible}\\

$\bullet$ Soaring oil costs and worries about global warming have prompted a search for \textbf{VIABLE} alternatives to fossil fuels. Some of the most \textbf{FEASIBLE} alternative energy sources include solar power, wind power, and biofuels. However, currently only around eight percent of energy in the United States comes from renewable sources, meaning that much research is still needed in order to find \textbf{VIABLE} alternative energy sources. Companies like BP and GE have invested billions of dollars in research on the most \textbf{FEASIBLE} sources of energy. 

\item \textbf{ANGUISH}\\
\textit{ Agonizing physical or mental pain; torment}\\

$\bullet$ The movie Batman Begins opens with a young boy's
\textbf{ANGUISH}. Eight-year-old Bruce Wayne falls into a
cave, where he encounters a swarm of bats. Bruce
develops a fear of bats and later urges his parents to
leave an opera featuring bat-like creatures. Outside
the theater, Bruce's parents are both killed in a
robbery. Filled with \textbf{ANGUISH}, Bruce blames himself
for his parents' murder and dedicates himself to
seeking revenge by fighting the criminals who control
Gotham City. As the Caped Crusader, Batman, Bruce
wages a successful fight against crime but must face
new and even more \textbf{ANGUISHING} questions: Does
his crusade have an end? Can he ever have an
ordinary life?

\item \textbf{INTEMPERATE}\\
\textit{ Lacking restraint; excessive }\\

\textbf{TEMPERATE}\\
\textit{ Exercising moderation and restraint}\\

$\bullet$ \textbf{INTEMPERATE} habits such as smoking, drinking, and overeating are \textbf{INIMICAL} (harmful) to good health. In contrast, a \textbf{TEMPERATE} person leads a lifestyle characterized by moderation and self-restraint. Bluto ( Animal House), Frank \say{The Tank} ( Old School), and Ben Stone ( Knocked Up) were all fun-loving, \textbf{INTEMPERATE} party animals. Compare their lifestyles to Andy Stitzer's ( The 40- Year-Old Virgin) far more \textbf{TEMPERATE} approach to life. The 18th century British author Samuel Johnson is famed for saying, \say{\textbf{ABSTINENCE} (refraining from use) is as easy to me as \textbf{TEMPERANCE} would be difficult.} 

\item \textbf{SUPERFICIAL}\\
\textit{ Shallow; lacking in depth; concerned with surface appearances}\\

 $\bullet$ What do Cher ( Clueless) and Daisy Buchanan ( The Great Gatsby) have in common? Both are \textbf{SUPERFICIAL}. In Clueless, Josh calls Cher \say{a \textbf{SUPERFICIAL} space cadet} because she lacks direction. Daisy proves to be a \textbf{SUPERFICIAL} person who prizes material possessions. For example, she bursts into tears when Gatsby shows her his collection of English dress shirts because she realizes that he has now become seriously wealthy. Tragically, Gatsby discovers that beneath. Daisy's \textbf{SUPERFICIAL} surface there is only more surface.
 
\begin{tcolorbox}
\begin{center}
 \textbf{KNOW YOUR ROOTS}\\
\line(1,0){250}\\
 \textbf{ LATIN PREFIX:  SUPER, SUPRA} - \textit{ over, above, greater in quality}
\end{center}
 
 \textbf{SUPERCILIOUS} - \textit{ overbearing, proud, haughty SUPERFICIAL on the surface, shallow }\\
 \textbf{SUPERLATIVE} - \textit{ the best, in the highest degree}\\ \textbf{SUPERNATURAL} - \textit{  above and beyond all nature}\\ \textbf{SUPERSEDE} - \textbf{ to take the place of, to SUPPLANT} 
\end{tcolorbox}  

 \item \textbf{LAUD, EXTOL, TOUT, ACCLAIM}\\
 \textit{  To praise; applaud}\\
 
  What do the Beach Boys' classic song \say{California Girls} and Katy Perry's hit \say{California Gurls} have in common? Both songs \textbf{EXTOL} the beauty of California girls. The Beach Boys acknowledge that they are \textbf{BEGUILED} (enticed, captivated) by the way southern girls talk. They \textbf{LAUD} east coast girls for being hip. However, this doesn't shake their \textbf{CONVICTION} (firm belief) that California girls are \say{the cutest girls in the world.} Needless to say, Katy Perry \textbf{CONCURS} (agrees) with the Beach Boys. She proudly \textbf{TOUTS} the beauty of California's \textbf{ACCLAIMED} golden coast. But that is not all. The California boys \say{break their necks} trying to sneak a peek at the \textbf{VOLUPTUOUS} (very sensual) \say{California gurls.} And who can blame them? According to Katy, \say{California gurls} are \say{unforgettable Daisy Dukes, bikinis on top.}  
  
\begin{tcolorbox}
\begin{center}
 \textbf{Tip for a Direct Hit}\\
\line(1,0){250}\\
\end{center}
   \textbf{LAUDS} is the morning church service in which psalms of praise to God are sung. Note that the word \textbf{APPLAUD} contains the root word \textbf{LAUD}. \textbf{LAUD} and its synonyms \textbf{EXTOL}, \textbf{TOUT}, and \textbf{ACCLAIM} all mean to praise.
\end{tcolorbox}  

 \item \textbf{DISMISSIVE}\\
 \textit{ Showing \textbf{INDIFFERENCE} (Word 10) or disregard; rejecting}\\
 
 $\bullet$ What do the artist Jackson Pollock, the author J.K. Rowling, and the reggae singer and rapper Sean Kingston have in common? All three had to overcome \textbf{DISMISSIVE} critics. Bewildered critics ridiculed Pollock, calling him \say{Jack the Dripper.} \textbf{INDIFFERENT} (Word 10) editors at numerous publishing houses rejected J.K. Rowling's story about a boy wizard named Harry Potter. And Sean Kingston almost quit the music industry after his idols Timbaland and Pharrell \textbf{DISMISSED} his early recordings. 
 
\begin{tcolorbox}
\begin{center}
 \textbf{  KNOW YOUR ROOTS}\\
\line(1,0){250}\\
\textbf{LATIN PREFIX: MITT/MISS} $ - $ \textit{ to send}\\
\end{center}
   \textbf{EMIT} $ - $ \textit{to send out}\\
    \textbf{SUBMIT} $ - $ \textit{to send under, yield, resign, surrender} \\ \textbf{TRANSMIT} $ - $ \textit{ to send across, communicate, convey}\\ \textbf{REMIT} $ - $ \textit{to send back, pay money, diminish in intensity}\\ \textbf{OMIT} $ - $ \textit{ to send by, pass by, neglect, leave out}\\ \textbf{ADMIT} $ - $ \textit{to send to, let in, confess, concede}\\ \textbf{COMMIT} $ - $ \textit{to send together, entrust, pledge, memorize}\\ \textbf{PERMIT} $ - $ \textit{ to send through, allow}\\
 \textbf{DISMISS} $ - $ \textit{to send away, discharge, put out of mind}\\ \textbf{REMISS} $ - $ \textit{ (adj.) negligent, lax, careless}\\ \textbf{REMITTANCE} $ - $ \textit{ a payment sent to pay a bill}\\ \textbf{MISSION} $ - $ \textit{ a duty one is sent to perform}\\ \textbf{MISSILE} $ - $ \textit{ something sent through the air}\\ \textbf{MISSIVE} $ - $ \textit{ a note sent by messenger}\\ \textbf{EMISSARY} $ - $ \textit{a messenger sent on a mission}
\end{tcolorbox}   
 
\item \textbf{DISPARAGE}\\
\textit{ To speak of in a slighting or disrespectful way; belittle}\\

$\bullet$ Did you see the movie Transformers: Revenge of the Fallen? What was your opinion? Does it deserve to be \textbf{LAUDED} (Word 91) or \textbf{DISPARAGED}? You might be surprised to learn that Megan Fox, the actress who played Mikaela Banes, \textbf{DISPARAGED} Transformers director Michael Bay for focusing more on special effects than on acting. Fox also blasted Bay, calling him a dictator \say{who wants to be like Hitler on his sets.} \textbf{GALLED} (irked) by Fox's \textbf{DISPARAGING} remarks, Bay shot back that Fox is young \say{and has a lot of growing to do.} Bay finally ended the war of words when he cut Fox from Transformers: Dark Of The Moon, saying her role was not \textbf{INTEGRAL} (essential) to the story 

\item \textbf{POMPOUS} \\
\textit{Filled with excessive self-importance; \textbf{PRETENTIOUS}; \textbf{OSTENTATIOUS} (Word 413); boastful}\\

$\bullet$ In the Harry Potter \textbf{SAGA} (Word 236), Draco Malfoy is a bully who arrogantly proclaims that pure - blood wizards are far superior to Muggles (non-wizards) and Mudbloods (Muggle-born witches and wizards). The \textbf{POMPOUS} Malfoy loves to use verbal taunts to \textbf{DENIGRATE} (malign) Harry, Ron, and Hermione. Draco is a literary \textbf{FOIL} (contrast) to the modest hero, Harry Potter. 

\item \textbf{CRYPTIC}\\
\textit{ Having a hidden or \textbf{AMBIGUOUS} (Word 21) meaning; mysterious}\\

$\bullet$ As Harry Potter and the Chamber of Secrets opens, Dobby delivers this \textbf{CRYPTIC} message to Harry: \say{Harry Potter must not go back to Hogwarts.} But why must Harry stay away from Hogwarts? Since the message is so \textbf{CRYPTIC}, we don't know. Later in the same book, a \textbf{CRYPTIC} message appears on one of the walls at Hogwarts: \say{The Chamber of Secrets has been opened. Enemies of the Heir, Beware.} Once again, since the message is \textbf{CRYPTIC}, we are not sure what it means. 

\item \textbf{SUBTLE}\\
\textit{ Difficult to detect; faint; mysterious; likely to elude perception}\\

$ \bullet $ Iago, the ultimate villain of English literature, is brilliantly \textbf{SUBTLE} in the way he manipulates Othello into believing that his wife, Desdemona, has been unfaithful. Iago \textbf{SUBTLY} plants suspicion with diversions, suggestions, and \textbf{INNUENDOES} (Word 66). This \textbf{SUBTLETY} makes Othello more deeply \textbf{APPREHENSIVE} (wary), and so Iago's \textbf{NEFARIOUS} (Word 139) plan succeeds in destroying both Othello and Desdemona.

\item \textbf{DISPARITY}\\
 \textit{An inequality; a gap; an imbalance}\\
 
$ \bullet $  The Hunger Games takes place in the nation of Panem, which contains 12 districts controlled by the \textbf{DESPOTIC} (Word 270) President Snow, who rules the country from the Capitol. There is great \textbf{DISPARITY} in Panem between the \textbf{AFFLUENT} (having a great deal of money, wealthy) and spoiled citizens of the Capitol and the \textbf{IMPECUNIOUS} (Word 138) residents of the districts, who live in \textbf{DEPLORABLE} (very bad) conditions and suffer from starvation. The \textbf{DISPARITY} is especially evident during the annual Hunger Games, in which the districts are forced to send teenagers to compete in a \textbf{MORTAL} (deadly) battle for the Capitol's entertainment. Mumbai (formerly Bombay) is India's financial capital and largest city. The movie Slumdog Millionaire features vivid images of the \textbf{DISPARITY} between the \textbf{AFFLUENT} (Word 257) few who live in the city's luxury condominiums and the poverty-stricken masses who live in tiny shacks in the densely-crowded Dharavi slum. 
\begin{tcolorbox}
\begin{center}
 \textbf{Tip for a Direct Hit}\\
\line(1,0){250}\\
\end{center}
\textbf{DISPARITY} contains the Latin root \textbf{PAR} meaning \say{that which is equal.} The root still lives in the golfing term \textbf{PAR}, which means to be equal to the course. It can also be seen in the SAT word \textbf{PARITY}, which means equality in status or value.  
\end{tcolorbox}  

\item \textbf{CURTAIL}\\
\textit{ To cut short or reduce}\\

$ \bullet $ The 2010 Gulf Oil Spill created an \textbf{UNPRECEDENTED} (Word 285) environmental and economic disaster. As a toxic oil slick spread across the Gulf's once \textbf{PRISTINE} (Word 417) beaches and wetlands, \textbf{IRATE} (angry, incensed) workers lost jobs while worried tourists \textbf{CURTAILED} and even canceled vacation trips to the region. The spill \textbf{UNDERSCORED} (emphasized) America's dependence upon gasoline. On average, Americans consume about 386 million gallons of gasoline each day. This \textbf{PRODIGIOUS} (huge) rate of consumption cannot go on forever. Many \textbf{PUNDITS} (Word 117) argue that Americans must \textbf{CURTAIL} their fuel consumption by developing renewable sources of energy.

\item \textbf{INNOCUOUS} \\
\textit{Harmless; unlikely to give offense or to arouse strong feelings or hostility; not \textbf{INIMICAL} }\\

$ \bullet $ Many mushrooms are \textbf{INNOCUOUS}, but there are some, like the Amanita or Death Cap mushroom, that are poisonous and should not be eaten. Sometimes a person will say something unkind and then claim that the intent was \textbf{INNOCUOUS}, saying, \say{Oh, they know I'm kidding.} Such an assertion may very well be \textbf{DISINGENUOUS} (Word 428), for the speaker is probably quite aware of the toxic effect of the not-so-\textbf{INNOCUOUS} words.

 \item \textbf{DIATRIBE, TIRADE, HARANGUE}\\
 \textit{ A bitter abusive denunciation; a thunderous verbal attack; a \textbf{RANT}} \\
 
 $ \bullet $ What do Coach Carter (Coach Carter), Coach Gaines (Friday Night Lights), and Coach Boone ( Remember the Titans) all have in common? All three coaches are passionate about building character and team-work. And, if necessary, all three don't hesitate to deliver a \textbf{TIRADE} when a player fails to follow team rules or perform to the best of his ability. For example, Coach Boone demands perfection. In one memorable \textbf{DIATRIBE} he insists, \say{We will be perfect in every aspect of the game. You drop a pass, you run a mile. You miss a blocking assignment, you run a mile. You fumble the football, and I will break my foot off in your John Brown hind parts and then you will run a mile. Perfection. Let's go to work!} It is debatable as to whether \textbf{HARANGUING} others in order to inspire them to different behaviours is an effective strategy. Former Indianapolis Colts' coach, Tony Dungy, refused to \textbf{RANT} at his players and achieved great success including winning the Superbowl.
\chapter{YOU MEET THE MOST INTERESTING
PEOPLE ON THE SAT: 101-130}

History is filled with a fascinating array of men and women who have both made enduring contributions and caused great tragedies. This chapter will introduce you to 30 SAT words that describe an astonishing variety of people. For example, you will meet Pharaoh Akhenaton, the ancient world's most famous ICONOCLAST, and Bill Gates, the modern world's most generous PHILANTHROPIST. As you study this chapter, you will learn words that will help you describe great orators, notorious traitors, and astute political commentators. That's why we are convinced that you meet the most interesting people on the SAT!

\item  \textbf{CHARLATAN}\\
\textit{Afake; fraud; imposter; cheat}\\

$ \bullet $Would you trust the Wizard of Oz, Gilderoy Lockhart (Harry Potter and the Chamber of Sec1·ets), or Frank Abagnale Jr. (Catch Me If You Can)? I hope not. All
three of these men were \textbf{CHARLATANS} or imposters who could not be trusted. The Wizard of Oz was a \textbf{CHARLATAN} who tried to trick Dorothy and her friends. Gilderoy Lockhart was a \textbf{CHARLATAN} who interviewed famous wizards and witches and then took credit for their heroic deeds. And Frank Abagnale Jr. was a \textbf{CHARLATAN} who pretended to be an airline pilot and a surgeon.

\begin{tcolorbox}
\begin{center}
\textbf{Tip for Direct Hit}\\
\line(1,0){250}\\
\end{center}
The word \textbf{CHARLATAN} often appears in sentence completion questions. Its important to remember that a \textbf{CHARLATAN} is associated with negative traits. A \textbf{CHARLATAN} will try to \textbf{DUPE} (mislead), \textbf{UNWARY} (incautious) victims with \textbf{SPURIOUS} (false) information.
\end{tcolorbox}

\item  \textbf{SKEPTIC}\\
\textit{A person who doubts; a skeptic asks questions
and lacks faith}

$ \bullet $ In the movie Men in Black, Edwards was originally a
\textbf{SKEPTIC} who did not believe that aliens were actually
living in NewYork City. In Bruce Almighty, Bruce was
originally a \textbf{SKEPTIC} who did not believe that the
man he met was really God. And in the movie Superbad, Seth was originally a \textbf{SKEPTIC} who did not
believe Fogell's fake ID, with the name \say{McLovin}
from Hawaii, would work.

 \item \textbf{RHETORICIAN}\\
\textit{An eloquent writer or speaker; a master of
RHETORIC (the art of speaking and writing)}

$ \bullet $ Frederick Douglass, Franklin Roosevelt, Martin
Luther King Jr., John F. Kennedy, and Ronald Reagan
were all \textbf{CHARISMATIC} (magnetic and inspidng)
leaders and superb \textbf{RHETORICIANS} whose eloquent
speeches inspired millions of people. For example, in
his inaugural address, President Kennedy challenged
Americans by proclaiming, \say{And so, my fellow
Americans: ask not what your country can do for you
ask what you can do for your country}.

 \item \textbf{HEDONIST}\\
\textit{A person who believes that pleasure is the chief
goal of life}

$ \bullet $ In Ancient Greece, the \textbf{HEDONISTS} urged their
followers to \say{eat, drink, and be merry, for tomorrow
we die}. Although it is a long way from Ancient Greece
to the home of rapper Ricky Ross in Miami, the
\textbf{HEDONISTIC} principle of pursuing pleasure remains
the same. During the tour of his \say{crib} Ross proudly
displayed the interior of his Escalade Maybach, a
Cadillac Escalade with the interior of a Maybach.
Hooked up with leather seats, plasmas, and satellites,
the interior provides everything a \textbf{HEDONIST} could
possibly ask for and more.

 \item \textbf{ASCETIC}\\
\textit{A person who gives up material comfmts and
leads a life of self-denial, especially as an act of
religious devotion}

$ \bullet $ At the age of 29, Prince Siddhartha Gautama left the
luxuries of his father's palace and for the next six
years adopted an extreme \textbf{ASCETIC} life. For days at a
time, he ate only a single grain of rice. His stomach
became so empty that, by poking a finger into it, he
could touch his backbone. Yet, Gautama gained only
pain, not wisdom. He decided to give up extreme
\textbf{ASCETICISM} and seek wisdom in other ways.
Gautama was successful and soon became known as
Buddha, a title meaning \say{the Enlightened One}.

\item \textbf{RACONTEUR}\\
\textit{A person who excels in telling \textbf{ANECDOTES}}

$ \bullet $ Herodotus was an ancient Greek historian who was a
renowned \textbf{RACONTEUR}. Many of the \textbf{ANECDOTES}
(Word 213) in the movie 300 are taken from his
famous history of the Persian Wars. For example,
Herodotus recounts how a Persian officer tried to
intimidate the Spartans by declaring that \say{A thousand nations of the Persian Empire descend upon you. Our arrows will blot out the sun}. \textbf{UNDAUNTED}
(Word 73), the Spartan warrior Stelios retorted, \say{Then
we will fight in the shade}.

\item\textbf{ICONOCLAST}\\
\textit{A person who attacks and ridicules cherished
figures, ideas, and institutions}

$ \bullet $ What do the Egyptian pharaoh Akhenaton and the
modern filmmaker Michael Moore have in common? Both are \textbf{ICONOCLASTS}. Akhenaton challenged
ancient Egypt's longstanding belief in a large number
of gods by rejecting polytheism and insisting that
Aton was the universal or only god. Michael Moore is
a modern \textbf{ICONOCLAST} whose documentary films
have attacked the Iraq War, the American health care
system, Wall Street bankers, and Washington
politicians. Like a true \textbf{ICONOCLAST}, Moore
ridiculed Congress, saying that most of its members
are scoundrels who deserve to be \say{removed and
replaced}.

\item\textbf{DILETTANTE}\\
\textit{An amateur or dabbler; a person with a
SUPERFICIAL (Word 90) interest in an art or
a branch of knowledge}

$ \bullet $ In the movie \say{Iron Man}, Tony Stark enjoys being a
\textbf{DILETTANTE} playboy who lets Obadiah Stane take
care of the day-to-day operations of Stark Industries.
However, behind his facade of being a \textbf{DILETTANTE}
playboy, Tony is in reality a \textbf{POLYMATH} (a person of
great and varied learning) who is a master engineer
and inventor. Held prisoner by a terrorist group, Tony
battles his way out of captivity by building a
\textbf{PROTOTYPE} (Word 34) armored suit.

\item\textbf{PARTISAN}\\
\textit{A person with strong and therefore biased
beliefs}

$ \bullet $ Are you pro-life or pro-choice? Do you favor staying
the course in Afghanistan or withdrawing the troops?
Do you think bailing out the big Wall Street banks was
a good or a bad policy? If you have a strong view on these issues, you are a \textbf{PARTISAN}. Remember a
\textbf{PARTISAN} speaks up and welcomes controversy. ln
contrast, \textbf{NONPARTISAN} issues enjoy widespread
public support. For example, during the Cold War,
most Americans supported the policy of containing
Soviet expansion.

\item \textbf{MENTOR}:
\textit{An advisor; teacher; guide}\\
\textbf{ACOLYTE}
\textit{A devotedfollower}

$ \bullet $ In the Star Wars \textbf{SAGA} (Word 208), Obi-Wan Kenobi is a Jedi Knight who served as Luke Skywalker's
\textbf{MENTOR}. As an eager young \textbf{ACOLYTE}, Skywalker learned the ways of the Force, a natural power
harnessed by the Jedi in their struggle against the
VILLAINOUS (vile, wicked) Darth Vader and evil
Galactic Empire.

\item\textbf{DEMAGOGUE}\\
\textit{A leader who appeals to the fears, emotions,
and prejudices of the populace}

$ \bullet $ Adolf Hitler is often cited as the epitome of a \textbf{DEMAGOGUE}. Hitler rose to power by using impassioned
speeches that appealed to the ethnic and nationalistic
prejudices of the German people. Hitler exploited,
embittered, and misled war veterans by blaming their
plight on minorities and other convenient scapegoats.
Unfortunately, Americans have not been immune to
the impassioned pleas of \textbf{DEMAGOGUES}. During the
1950s, Senator Joseph McCarthy falsely alleged that
Communist sympathizers had infiltrated the State Department. As McCarthy's \textbf{DEMAGOGIC} rhetoric
grew bolder, he \textbf{DENOUNCED} (Word 176) General
George Marshall, former Army Chief of Staff and ex Secretary of State, as \say{part of a conspiracy so immense
and an infamy so black as to dwarf any previous
venture in the history of man}.

\item \textbf{AUTOMATON}\\
\textit{A self operating machine; a mindless follower;
a person who acts in a mechanical fashion}

$ \bullet $ In the Harry Potter series, the Imperius Curse was a
spell that caused its victim to fall under the command
of the caster. In Hary Potter and the Deathly
Hallows, the Death Eater Yaxley placed an Imperius
Curse on Pius Thickness. When Thickness became
Minister of Magic, be behaved like an \textbf{AUTOMATON}
or mindless follower of Lord Voldemort.

\item \textbf{RECLUSE}\\
\textit{A person who leads a secluded or solitary life}

$ \bullet $ What do Harper Lee, Sybill Trelawney, and Greta
Garbo have in common? All three were \textbf{RECLUSES}
who wanted to live alone. Although she is the world famous author of To Kill a Mockingbird, Harper Lee
rarely appears in public. Sybill Trelawney was the
Divination professor at Hogwarts who lived alone in
the North Tower because she didn't want to \say{cloud her
Inner Eye}. And Greta Garbo was a famous actress
who summed up what it means to be a \textbf{RECLUSE}
when she said: \say{I want to be alone}.

\item\textbf{BUNGLER}\\
\textit{Someone who is clumsy or \textbf{INEPT}; a person
who makes mistakes because of incompetence}

$ \bullet $ \textbf{BUNGLERS} have been featured in a number of
movies and television programs. For example, The
Three Stooges were a trio of \textbf{BUNGLERS} whose
\textbf{INEPT} blunders and madcap antics never failed to
leave their fans laughing. In the movie Mighty Ducks,
the team was a group of \textbf{BUNGLERS} that did not
know how to play hockey or work together. One of
television's most beloved .\textbf{BUNGLERS} was Gilligan,
the clumsy first mate on Gilligan's Island.

\item\textbf{ CLAIRVOYANT}\\
\textit{Having the supposed power to see objects and
events that cannot be perceived with the.five
traditional senses; a \textbf{SEER}}

$ \bullet $ Sybill Trelawneywas the Divination professor at Hogwarts who claimed to be a \textbf{CLAIRVOYANT}. She used
tea leaves and crystal balls to see the future. Both
Harry and Professor Dumbledore were \textbf{SKEPTICAL}
(Word 102) about her claim to be a \textbf{CLAIRVOYANT}.
In Harry Potter and the Order of the Phoenix,
Dolores Umbridge fired Sybill for being a \textbf{CHARLATAN} (Word 101). Nonetheless, readers ofthe Harry
Potter series know that Trelawney did make two
extremely important and very accurate prophecies.

 \item \textbf{PROGNOSTICATOR}\\
\textit{A person who makes predictions based upon
current information and data}

$ \bullet $ Weather forecasters, sports announcers, and financial
analysts are all \textbf{PROGNOSTICATORS} who use information and data to make predictions and forecasts. It is
important to understand the difference between a
\textbf{PROGNOSTICATOR} and a \textbf{CLAIRVOYANT} (Word 115).
Although both make predictions, a \textbf{PROGNOSTICATOR} uses empirical data that can be collected, seen,
and studied. In contrast, a \textbf{CLAIRVOYANT} claims to
see the future through means beyond the five senses.

 \item. \textbf{PUNDIT}\\
\textit{An expert commentator; an authority who
expresses his or her opinion, usually on
political issues}

$ \bullet $ From CNN's News Center to ESPN's Sports Center,
television programs are filled with \textbf{PUNDITS} who
offer their "expert" commentary on issues ranging
from political campaigns to March Madness brackets.
The \textbf{PUNDITS} almost always sound authoritative and
convincing. But it is wise to maintain a healthy
\textbf{SKEPTICISM} (Word 102). Here are expert opinions
from famous pundits that turned out to be wrong:\\

\say{\textit{Louis Pasteur's theory of germs is ridiculous
fiction}}.\\
Pierre Packet, Professor of Physiology at
Toulouse, 1872\\

\say{\textit{Heavier-than-air flying machines are
impossible}}.\\
Lord Kelvin, President of the Royal Society,
1895\\

\say{\textit{Stocks have reached what looks like a
permanently high plateau}}.\\
Irving Fisher, Professor of Economics, Yale
University, 1929.\\

 \say{\textit{There is no reason anyone would want a
computer in their home}}.\\
Ken Olson, President, Chairman, and
Founder of Digital Equipment Corp, 1977

 \item \textbf{ZEALOT}\\
\textit{A very enthusiastic person; a champion; a true
believer}

$ \bullet $ William Lloyd Garrison was a \textbf{ZEALOT} who championed the cause of unconditional and immediate
abolition. In the first issue of The Liberator, Garrison
left no doubt as to his intentions when he wrote: \say{I am
in earnest - I will not equivocate - I will not excuse -
I will not retreat a single inch - AND I WILL BE
HEARD}.
 \item \textbf{NEOPHYTE, NOVICE, GREENHORN}\\
\textit{All three are beginners}

$ \bullet $ In October 2008, Justin Bieber was an unknown
\textbf{NEOPHYTE} who had never professionally recorded a
song. However, Usher recognized that although
Bieber was \textbf{NOVICE}, he was a musical \textbf{PRODIGY}
(Word 123) who had the potential to become a
superstar. With Usher as his \textbf{MENTOR} (Word 110),
the angelic-looking Bieber soon developed a \say{streetwise} look that included baseball caps, hoodies, hip
hop chains, and flashy sneakers. Usher quickly
transformed Bieber from a \textbf{GREENHORN} to a global
sensation. In July 2010, JB's music video Baby
\textbf{SUPPLANTED} (replaced) Lady Gaga's Bad Romance
video as the most viewed YouTube video ever.

 \item \textbf{BENEFACTOR and PATRON}\\
\textit{A person who makes a gift or bequest}\\
\textbf{BENEFICIARY}\\
\textit{The recipient of funds, titles property, and
other benefits}

$ \bullet $ Nicholas Sparks has achieved international fame by
writing romance novels such as The Notebook and A
Walk to Remember· that are often set in New Bern,
North Carolina. Residents of New Bern also know
Sparks as a generous \textbf{BENEFACTOR} and \textbf{PATRON}
who has donated nearly \$1 million dollars to build a
state-of-the-art track and field facility for New Bern
High School. As the \textbf{BENEFICIARIES} of this \textbf{MUNIFICENT} (Word 232) gift the New Bern Bears have
become one of North Carolina's top track and field
teams. Note that both \textbf{BENEFACTOR} and \textbf{BENEFICIARY} begin with the Latin prefix bene, which means \say{good.} So a \textbf{BENEFACTOR}, like Nicholas Sparks,
gives good gifts, and a \textbf{BENEFICIARY}, like New Bern
High School, receives good gifts.

 \item \textbf{DISSEMBLER and PREVARICATOR}\\
\textit{Both are liars and deceivers}

$ \bullet $ In Mean Girls, Regina George was a cunning \textbf{DISSEMBLER} who deliberately lied to her friends and to
her enemies. In the movie Pirates of the Caribbean:
Curse of the Black Pearl, Captain Barbarossa was a
\textbf{PREVARICATOR} who repeatedly lied to Jack Sparrow,
Elizabeth Swann, and Will Turner.


 \item\textbf{PROPONENT}\\
\textit{One who argues in support of something; an
\textbf{ADVOCATE}; a champion of a cause}

$ \bullet $ Although America has faced a number of challenging
social problems, our nation has always produced
leaders who were strong \textbf{PROPONENTS} of reform .
For example, during the 19th century, Jane Addams
was an outspoken PROPONENT for urban settlement
houses. Today, former Vice-President Al Gore is a
vigorous \textbf{ADVOCATE} of implementing measures that
will reduce global warming. One way to remember
\textbf{PROPONENT} is to note that the prefix pro means to
be for something.

\item \textbf{PRODIGY}
\textit{A person with great talent; a young genius}

$ \bullet $ What do Wolfgang Mozart and Pablo Picasso have in
common? Both were \textbf{PRODIGIES} who demonstrated
uncanny artistic talent at a young age. Mozart was a
child \textbf{PRODIGY} who wrote his first symphony at the
age of eight and grew into a \textbf{PROLIFIC} (Word 347)
adult who wrote over 6oo pieces of music before
dying at the age of 35. Like Mozart, Picasso
demonstrated \textbf{PRECOCIOUS} (very advanced) talent
by drawing pictures before he could talk. Picasso
mastered many styles but is best known as the
\textbf{PROGENITOR} (originator) of Cubism.

 \item\textbf{ORACLE}\\ 
\textit{A person considered to be a source of wise counsel or prophetic opinions}

$ \bullet $ Would you like to know what is going to happen in the
future? All you have to do is ask an \textbf{ORACLE}. While

the ancient Greeks asked the Delphic Oracle to predict
the future, World Cup soccer fans watched televised
reports featuring the predictions of an octopus named
Paul. The eight-legged \textbf{ORACLE} became a global
sensation when be correctly predicted the winner of
eight straight matches. Paul's \textbf{PROGNOSTICATIONS}
(Word 116) have attracted \textbf{LUCRATIVE} (Word 227)
offers from people who want to know the outcome of
elections and the gender of future children.

\item \textbf{MISANTHROPE}\\
\textit{A person who hates or distrusts humankind}

$ \bullet $ Ebenezer Scrooge and Alceste are two of the best
knovm \textbf{MISANTHROPES} in literature. Scrooge is the
main character in Charles Dickens's 1843 novel, A
Christmas Carol. He is a cold-hearted, \textbf{MISERLY}
(very stingy) \textbf{MISANTHROPE} who despises poor
people and Christmas. Alceste is the main character in
Moliere's 1666 play, Misanthrope. He is a judgmental
\textbf{MISANTHROPE}, quick to criticize the flaws in people.

 \begin{tcolorbox}
\begin{center}
\textbf{Tip for Direct Hit}\\
\line(1,0){250}\\
\end{center}
\textbf{MISANTHROPE} combines the Greek prefix
miso meaning \say{hate} with the Greek root anthropos meaning \say{humankind} Prefixes make a difference in the meaning of words. If we place
the Greek prefix philo, meaning \say{love} in front of anthropos, we will have the word \textbf{PHILANTHROPY}, meaning love of humankind. A \textbf{PHILANTHROPIST} loves humanity so much that he or she donates time and money to charity.
\end{tcolorbox}


\item \textbf{INNOVATOR}\\
\textit{A person who introduces something new}

$ \bullet $ What do D.W. Griffith and James Cameron have in
common? Both are cinematic \textbf{INNOVATORS} whose
pioneering works have had a profound impact on
movies. Griffith's groundbreaking film The Birth of a
Nation (1915) included such \textbf{SEMINAL} (highly
influential and original) camera techniques as jump-cuts and facial close-ups. Running an
\textbf{UNPRECEDENTED} (Word 259) 3 hours and 10
minutes, The Birth ofa Nation ushered in a new era
of blockbuster films.

$ \bullet $ Like D. W. Griffith, James Cameron is recognized as
an \textbf{INNOVATOR} whose work is redefining the art of
special effects. Avatar (2009) is a visually stunning
movie that uses a 3D Fusion Camera System to
seamlessly sew human actors into a digital world in
real time.

\item \textbf{SYCOPHANT}\\
\textit{A person who seeks favor by flattering people
of influence; a \textbf{TOADY}; someone who behaves
in an \textbf{OBSEQUIOUS} or servile manner}

$ \bullet $ Louis XIV compelled France's great nobles to live at
the Versailles Palace. Life at the royal palace
transformed \textbf{HAUGHTY} (arrogant) aristocrats into
favor-seeking \textbf{SYCOPHANTS}. Instead of competing
for political power, nobles \textbf{SQUANDERED} (wasted)
their fortunes jockeying for social prestige. For
example, nobles vied for the \textbf{COVETED} (Word 32)
honor of holding Louis XIV's shirt as he prepared to
get dressed.

\item \textbf{STOIC}\\
\textit{A person who is seemingly \textbf{INDIFFERENT}
(Word 10) to or unaffected by joy, grief,
pleasure, or  pain; someone who is impassive
and emotionless}\\

$ \bullet $ What would you do if you scored the winning goal in a
championship soccer game? What would you do if
your error caused your team to lose a championship
baseball game? Most people would be elated to win
and dejected to lose. However, a \textbf{STOIC} would remain
impassive, showing no emotion in victory or defeat.

$ \bullet $ Being a \textbf{STOIC} is not easy. It requires great discipline
and self-control. For example, tourists to London are
familiar with the distinctive bearskin helmets and
scarlet uniforms worn by the guards at Buckingham
Palace. The guards are famous for their ability to
\textbf{STOICALLY} endure hot summer weather and hordes
of pesky tourists.

\item \textbf{REPROBATE}\\
\textit{A morally unprincipled person}

$ \bullet $ Who is the most despised \textbf{REPROBATE} living in
America today? For thousands of betrayed investors
there is only one answer- Bernard Madoff. On June
29, 2009, Judge Denny Chin sentenced Madoff to 150
years in prison for running a giant Ponzj scheme that
cheated investors out of almost 65 billion dollars.
Madoffs victims included pension funds, charitable
institutions, and elderly retirees. Although Madoff
was a \textbf{CHARLATAN} (Word 101), he is best described
as a \textbf{REPROBATE} because of the enormity of a fraud
that Judge Chin called \say{extraordinarily evil}.

\item \textbf{RENEGADE}\\
\textit{A disloyal person who betrays his or her cause; a traitor; a deserter}

$ \bullet $ In 1777, Benedict Arnold was one of America's most
admired Revolutionary War generals. Yet, just three
years later, Arnold was vilified as a \textbf{RENEGADE}
whose name became synonymous with traitor. What
happened to cause this amazing change in Arnold's
reputation? Despite his bravery at the pivotal battle of
Saratoga, Arnold was passed over for promotion while
other officers took credit for his accomplishments.
Frustrated and bitter, Arnold secretly became a
British agent. In 1780, he obtained command of West
Point in order to surrender it to the British. American
forces discovered Arnold's treacherous scheme, and
he was forced to flee to London to avoid capture.
Today, Arnold's contributions to the colonial cause
are forgotten, and he is remembered as our nation's
first and foremost \textbf{RENEGADE}.

\begin{tcolorbox}
\begin{center}
\textbf{Tip for Direct Hit}\\
\line(1,0){250}\\
\end{center}
The words \textbf{REPROBATE} (Word 129) and \textbf{RENEGADE} (Word 130) are easy to confuse. They sound similar and both are negative words used to describe despicable people. Nonetheless, there are important differences between the people describe by these two words. A  \textbf{REPROBATE} is the best remembered as a morally unprincipled and evil person. A  \textbf{RENEGADE} is the best remembered as a traitor and deserter.
\end{tcolorbox}



\end{enumerate}
\end{document}