\documentclass[a4paper,11pt]{article}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{evan}
\usepackage{lmodern}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{hyperref}
\usepackage{graphicx}
\usepackage[english]{babel}
\usepackage{amsmath}
\usepackage{dirtytalk}
\usepackage{mathtools}
\usepackage{tikz}
\usetikzlibrary{arrows,positioning} 
\usepackage{array}
\usepackage{wrapfig}
\usepackage{multirow}
\usepackage{tabu}
\usetikzlibrary{calc}
\usepackage{changepage}
\usepackage{caption,setspace}
\usepackage{draftwatermark}
\newcommand\perm[2][^n]{\prescript{#1\mkern-2.5mu}{}P_{#2}}
\newcommand\comb[2][^n]{\prescript{#1\mkern-0.5mu}{}C_{#2}}
\SetWatermarkText{ $Rejected! $}
\begin{document}
"Pardon me if I tell you that I don't miss my BUET life at all. In fact it could hardly be a life beyond survival. My story started from my college life in Rajshahi. It was like a miracle that I became champion in division level of Math Olympiad and then again in national level. Consequently, I was selected for the math camp and was preparing for IMO with eyes full of dreams. 6 people get selected every year but unfortunately I missed that chance with having all the qualities but lacking in previous experience. That was the first time my luck betrayed me. But my camp experience made me more confident than I was ever. I came back to Rajshahi and concentrated on HSC studies and was taking preparations for SAT parallely. I applied to Brown University and Colby College in USA but I also knew that in any case I shouldn't miss BUET. And I was able to reach there finally. I started my class. Then it was almost midterm, I went to Singapore to sit for the admission test of NTU. When I was in Singapore, I received a call and got to know that I was selected to represent team Bangladesh in International Physics Olympiad in Thailand. I came back to Dhaka. It was a clumsy situation. If I go to Thailand then I would be going to miss my term finals. I talked to my advisor and he was like what's the point in participating in that contest! I was hurt and speechless. Then I went to the office of VC. He heard my story told me that he would give me a chance of withdrawing my courses for that term and advised me to wait for a year and then start classes with the next batch. Same thing was suggested from the register office. But I was misinformed. I could actually continue with my own batch and then re-registered for my remaining courses. But that one mistake changed my life completely. That time I took the risk because I thought it was worth. Why do we always settle for less than we deserve? Actually the decision of participating in Physics Olympiad was the toughest decision of my life. I didn't know what had been waiting for me in future. I withdrew my courses from BUET and was rejected by almost all the foreign colleges I applied. National University of Singapore (NUS) put me on the waiting list. Finally they approved me an admission without funding. If I continued there I would have to do an additional part time job to bear my expense. I decided that I wouldn't go for it. May be that time in my life I overestimated my studies and wasn't ready to make any compromise for it. The next year I discovered me in a class where I would never be able to belong in coming four years. This was going to be the only class room that had two roll 21s, one from batch 10 and another from 11, throughout the sessions. It was hard to bear this complex identity. I had to face the eyes of mistrust of my teachers if I got good marks. I had to explain myself all the time but why would I? I knew that there are some judgements that everyone in my class room were sure of, but no one actually said. I shared this place for a long period of time with people who couldn't give me the acceptance that I longed for in my heart. I am not a lagger, trust me guys! The more I expected from people's response to my experience of abandonment, the more I was feeling exposed to ridicule. Imagine that someone left the seat beside you during a class test because you are not so useful! It was my mistake that I didn't allow myself to be a part of Batch 10 and it was tough for batch 11 to accept me. I was stuck somewhere between this batch and that batch. BUET is a campus you know where a senior is a senior. At last my classmates found me as a person they might potentially get along with. But it was too late. I was someone who was rejected by everything I loved. The man who loved mathematics too much, his highest grade in that subject was an A-. I lived my whole BUET life like a drop out. The system here made me feel suffocated all the time. But things always change with time. At one stage of my life I had dreams but now I have the courage to accept the alternate reality. I am now prepared to see what I don't want to. We exist in a universe that is held in complete balance. Can you really hear my voice when I'm talking about something very upsetting but trying to be strong? Life has made me so lonely. My fishes and plants were the best friends in my darkest times in Nazrul Islam hall. I helped them to grow. I discovered that fishes eat eggs of mosquitoes to live. It made me realize that every single living being has a purpose and they are useful. And among all my plants, I liked the cactus most. You may find it weird. But you know what? Even a cactus can grow flowers."\\
.\hfill -Mahadi Masnad 


\end{document}